---
title: "Simulation with the Smallest Lambda for IpaB IgG"
format: 
  pdf:
    number-sections: true
    number-depth: 2
    number-offset: [0, 0]
    pdf-engine: pdflatex
header-includes: 
  - \usepackage{graphicx}  # Required for ggplot figures
  - \usepackage{float}  # Helps place figures properly
  - \usepackage{wrapfig}  # Now safely included since you installed it
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE,echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
library(tibble)
library(dplyr)
library(serocalculator)
library(haven)
library(knitr)
library(plotly)
library(kableExtra)
library(tidyr)
library(arsenal)
library(forcats)
library(huxtable)
library(magrittr)
library(parameters)
library(ggplot2)
library(ggeasy)
library(scales)
library(plotly)
library(patchwork)
library(tidyverse)
library(readxl)
library(purrr)
library(here)
library(table1)
library(furrr)
library(future)
```


# Vary simulation lambdas across the range in the real cross-sectional data

## 1. Estimate the shigella incidence rate using the real cross-sectional Shigella data, 
```{r,echo=FALSE, message=FALSE, warning=FALSE,results='hide'}
## Load data
load("~/ipab.RData")
```

## Get parameters from longitudinal data

```{r,echo=FALSE, message=FALSE, warning=FALSE,results='hide'}
## Set noise 
create_noise_df <- function(country) {
  noise_df <- tibble(
    antigen_iso = c("IgG"),
    Country     = factor(c(country)),  # Use input argument for Country
    y.low       = c(25),
    eps         = c(0.25),
    nu          = c(0.5),
    y.high      = c(200000)
  )
  
  return(noise_df)
}

noise_df_USA <- create_noise_df("MA USA")
noise_df_Niger <- create_noise_df("Niger")
noise_df_Sierra <- create_noise_df("Sierra Leone")
noise_df_Ghana <- create_noise_df("Ghana")
```

## 2. Run simulations using: 
```{r,echo=FALSE, message=FALSE, warning=FALSE}
# MA USA
est_USA <- est.incidence(
  pop_data = df_xs_USA_ipab_IgG,
  curve_params = curve_params_shigella,
  noise_params = noise_df_USA,
  antigen_isos = c("IgG"),
)

# Ghana
est_Ghana <- est.incidence(
  pop_data = df_xs_Ghana_ipab_IgG,
  curve_params = curve_params_shigella,
  noise_params = noise_df_Ghana,
  antigen_isos = c("IgG"),
)

# Niger
est_Niger <- est.incidence(
  pop_data = df_xs_Niger_ipab_IgG,
  curve_params = curve_params_shigella,
  noise_params = noise_df_Niger,
  antigen_isos = c("IgG"),
)

# Sierra Leone
est_Sierra <- est.incidence(
  pop_data = df_xs_Sierra_ipab_IgG,
  curve_params = curve_params_shigella,
  noise_params = noise_df_Sierra,
  antigen_isos = c("IgG"),
)

# create table of incidence.rate of each region

create_incidence_table <- function(...) {
  # Capture input objects and their names
  est_list <- list(...)
  country_names <- names(est_list)
  
  # Extract incidence.rate from the summary() of each estimate object
  incidence_rates <- sapply(est_list, function(x) summary(x)$incidence.rate)
  
  # Create a tidy tibble
  incidence_table <- tibble(
    Country = country_names,
    Incidence_Rate = incidence_rates
  )
  
  return(incidence_table)
}
# Example usage with four different country estimates:
incidence_summary <- create_incidence_table(
  USA = est_USA,
  Ghana = est_Ghana,
  Niger = est_Niger,
  Sierra_Leone = est_Sierra
)

# Display the summary table
print(incidence_summary)

# 1) The median of the incidence rate estimates from step 1.
# Get incidence rate estimates from each region (4 regions) and do median
med_ipab_IgG<-median(incidence_summary$Incidence_Rate)

#1.188

# 2) 2x the maximum incidence rate estimate from step 1
# Get incidence rate estimates from each region (4 regions) and get the maximum one and 2x
max2_ipab_IgG<-2*max(incidence_summary$Incidence_Rate)

# 2.788

# 3) 1/2 the minimum incidence rate from step 1
# Get incidence rate estimates from each region (4 regions) and get the minimum one and 1/2
min_half_ipab_IgG<-0.5*min(incidence_summary$Incidence_Rate)

#0.185
# Create a data frame for the incidence rate summary
incidence_rate_summary <- data.frame(
  Metric = c("Median Incidence Rate", 
             "2x Maximum Incidence Rate", 
             "1/2 Minimum Incidence Rate"),
  Value = c(median(incidence_summary$Incidence_Rate),  # Median of incidence rates
            2 * max(incidence_summary$Incidence_Rate),  # 2x Maximum incidence rate
            0.5 * min(incidence_summary$Incidence_Rate)) # 1/2 Minimum incidence rate
)

# Print the table
print(incidence_rate_summary)

```
These three values are preliminary observations of lambda from ipab_IgG.

We choose a lambda that is half of the minimum lambda from the four regions.
 
\newpage 

# Simulation (300times) with the smallest lambda
```{r,echo=FALSE, message=FALSE, warning=FALSE,results='hide'}

# Define the simulation function
simulate_seroincidence <- function(nrep, n_sim, observed, range = NULL) {
  # Set parallel plan inside function to avoid issues with distributed nodes
  plan(multicore)  # Use multiple cores for parallel processing (works best on HPC)
  
  # Parameters
  dmcmc <- curve_params_shigella  # Curve parameters
  antibodies <- c("IgG")  # Antigen-isotypes
  lambda <- observed  # Simulated incidence rate per person-year
  
  # Biologic noise distribution
  dlims <- rbind("IgG" = c(min = 0, max = 0.5))
  
  # Noise parameters
  cond <- tibble(
    antigen_iso = c("IgG"),
    nu = c(0.5),  # Biologic noise (nu)
    eps = c(0.25),  # Measurement noise (eps)
    y.low = c(25),  # Low cutoff (llod)
    y.high = c(200000)  # High cutoff (y.high)
  )
  
  # Perform simulations in parallel
  results <- future_map(1:n_sim, function(i) {
    tryCatch({
      # Generate cross-sectional data
      csdata <- sim.cs(
        curve_params = dmcmc,
        lambda = lambda,
        n.smpl = nrep,
        age_range = range,
        antigen_isos = antibodies,
        n.mc = 0,
        renew_params = TRUE,  # Use different parameters for each simulation
        add.noise = TRUE,
        noise_limits = dlims,
        format = "long"
      )
      
      # Estimate seroincidence
      est <- est.incidence(
        pop_data = csdata,
        curve_params = dmcmc,
        noise_params = cond,
        lambda_start = 0.1,
        build_graph = TRUE,
        verbose = FALSE,
        print_graph = FALSE,
        antigen_isos = antibodies
      )
      
      # Return results for this simulation
      list(csdata = csdata, est1 = est)
    }, error = function(e) {
      return(list(error = e$message))  # Capture and store errors instead of stopping execution
    })
  }, .options = furrr_options(seed = TRUE))
  
  # Ensure sequential processing after function execution
  plan(sequential)
  
  return(results)
}

# ------------------------ #
# 🔹 Run Simulations in Parallel
# ------------------------ #

# Set up parallel processing (Ensure future::plan() is set before execution)
plan(multicore)  # Works best on HPC

# Define parameter sets
sim_params <- list(
  list(nrep = 100, range = c(0, 2)),
  list(nrep = 100, range = c(2, 5)),
  list(nrep = 200, range = c(0, 2)),
  list(nrep = 200, range = c(2, 5)),
  list(nrep = 300, range = c(0, 2)),
  list(nrep = 300, range = c(2, 5)),
  list(nrep = 400, range = c(0, 2)),
  list(nrep = 400, range = c(2, 5))
)

# Run simulations in a loop to reduce redundant code
results_list <- lapply(sim_params, function(params) {
  cat("Running simulation for nrep =", params$nrep, "range =", params$range, "\n")
  simulate_seroincidence(nrep = params$nrep, n_sim = 300, observed = min_half_ipab_IgG, range = params$range)
})

# Assign results to variables
results_100_1 <- results_list[[1]]
results_100_2 <- results_list[[2]]
results_200_1 <- results_list[[3]]
results_200_2 <- results_list[[4]]
results_300_1 <- results_list[[5]]
results_300_2 <- results_list[[6]]
results_400_1 <- results_list[[7]]
results_400_2 <- results_list[[8]]

# Stop parallel processing
plan(sequential)  # Return to sequential processing after execution

```


## Store each sample size in table
```{r,echo=FALSE, message=FALSE, warning=FALSE,results='hide'}
# Define a function to generate final tables
generate_final_table <- function(results_list, sample_size) {
  # Initialize an empty list to store the results
  summary_results <- list()
  
  # Loop through each of the 100 results and extract the required columns
  for (i in 1:300) {
    # Extract the summary for each result
    result_summary <- summary(results_list[[i]]$est1)
    
    # Select the required columns
    extracted_columns <- result_summary %>%
      select(incidence.rate, SE, CI.lwr, CI.upr)
    
    # Add a column for the index (optional, for tracking)
    extracted_columns <- extracted_columns %>%
      mutate(index = i)
    
    # Append to the list
    summary_results[[i]] <- extracted_columns
  }
  
  # Combine all results into a single data frame
  final_table <- bind_rows(summary_results) %>%
    mutate(sample_size = sample_size) # Add sample size column for clarity
  
  return(final_table)
}

# Example usage for different sample sizes
final_table_100_1 <- generate_final_table(results_100_1, 100)
final_table_100_2 <- generate_final_table(results_100_2, 100)

final_table_200_1 <- generate_final_table(results_200_1, 200)
final_table_200_2 <- generate_final_table(results_200_2, 200)

final_table_300_1 <- generate_final_table(results_300_1, 300)
final_table_300_2 <- generate_final_table(results_300_2, 300)

final_table_400_1 <- generate_final_table(results_400_1, 400)
final_table_400_2 <- generate_final_table(results_400_2, 400)
```



## Graphs of where the x axis is the sample size and the y axis is the empirical standard error

Set the preliminary observed lambda as half of the minimum incidence rate from the four regions.

```{r,echo=FALSE, message=FALSE, warning=FALSE}
# Function to calculate metrics for each table
calculate_metrics <- function(data, sample_size) {
  empirical_se <- sd(data$incidence.rate)
  
  
  # Return a data frame with results
  data.frame(
    sample_size = sample_size,
    empirical_se = empirical_se
    
  )
}

# Apply the function to each table
metrics_100_1 <- calculate_metrics(final_table_100_1, 100)
metrics_100_2 <- calculate_metrics(final_table_100_2, 100)

metrics_200_1 <- calculate_metrics(final_table_200_1, 200)
metrics_200_2 <- calculate_metrics(final_table_200_2, 200)

metrics_300_1 <- calculate_metrics(final_table_300_1, 300)
metrics_300_2 <- calculate_metrics(final_table_300_2, 300)

metrics_400_1 <- calculate_metrics(final_table_400_1, 400)
metrics_400_2 <- calculate_metrics(final_table_400_2, 400)


# Combine the results into a single summary table
summary_metrics_1 <- bind_rows(metrics_100_1, metrics_200_1, metrics_300_1,
                               metrics_400_1)


summary_metrics_2 <- bind_rows(metrics_100_2, metrics_200_2, metrics_300_2,
                               metrics_400_2)

# Add a column to distinguish the datasets
summary_metrics_1 <- summary_metrics_1 %>%
  mutate(Age_Group = "Age 0-2")

summary_metrics_2 <- summary_metrics_2 %>%
  mutate(Age_Group = "Age 2-5")

# Combine both datasets into one
summary_metrics_combined <- bind_rows(summary_metrics_1, summary_metrics_2)

# Plot with color to differentiate age groups
ggplot(summary_metrics_combined, aes(x = sample_size, y = empirical_se, color = Age_Group)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Empirical Standard Error vs. Sample Size",
    x = "Sample Size",
    y = "Empirical Standard Error",
    color = "Age Group"
  ) +
  theme_minimal()
```

\newpage

## Table

```{r,echo=FALSE, message=FALSE, warning=FALSE}
summary_metrics_combined %>%
  kable()
```