---
title: "Chapter 1 — Main-text Figures & Tables"
subtitle: "Reproduces every main-text table & figure from library(shigella) + the regenerated .rda"
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    toc-location: right
editor: visual
---

```{=html}
<!--
PURPOSE
  GitHub / package-phase reproducibility document. With the shigella package
  installed (devtools::install() or devtools::load_all()), this renders every
  main-text table and figure straight from library(shigella) — no source() of R/.

HOW TO RUN
  1. Regenerate the fitted .rda once: run data-raw/00_build_case_data.R, then
     data-raw/01..05_*.R (these write the .rda into your manuscript_data_dir).
  2. Edit the CONFIG chunk: point `rda_dir` at that folder.
  3. (Table 1 + Figure 1 only) set the three raw-Excel paths; leave them unset to
     skip those two chunks — every other item builds from .rda alone.
  4. Render (Quarto) or run chunk-by-chunk.
-->
```

```{r}
#| label: config
#| message: false

## ---- EDIT THESE ----------------------------------------------------------
rda_dir <- "~/Data/Manuscript"   # folder holding the regenerated *.rda

## (Table 1 + Figure 1 only) raw Excel inputs. Leave as-is to skip those two.
raw_compiled <- "~/path/to/3.8.2024 Compiled Shigella datav2.xlsx";  sheet_compiled <- "Compiled"
raw_metadata <- "~/path/to/Additional metadata for Luminex sample set 10.28.2024.xlsx";  sheet_metadata <- "Metadata"
raw_durdia   <- "~/path/to/Duration of diarrhea prior to presentation.xlsx";     sheet_durdia   <- "Duration of symptoms"
## --------------------------------------------------------------------------
```

```{r}
#| label: setup
#| message: false
#| warning: false

# Everything comes from the installed package: shigella's Imports (dplyr, tidyr,
# purrr, ggplot2, patchwork, scales, flextable, gt, cowplot, serocalculator,
# serodynamics, ...) load with it, and the flextable/gt knit_print methods
# register on load. No source() of R/ is needed.
library(shigella)
```

```{r}
#| label: load-rda
#| message: false

# Every object the main-text figures/tables need, loaded from rda_dir.
main_rda <- c(
  # overall fits
  "overall_IpaB_pop_6", "overall_Sf2a_pop_6", "overall_Sf3a_pop_6",
  "overall_Sf6_pop_6",  "overall_Sonnei_pop_6",
  # serotype-specific fits
  "serotype_sf2a_3", "serotype_sf3a_3", "serotype_sonnei_3",
  # combined flexneri fits
  "combined_flexneri_sf2a", "combined_flexneri_sf3a",
  # IpaB age fits (Table 3 + Fig 5A)
  "overall_IpaB_pop_under5", "overall_IpaB_pop_plus5",
  # case data, DurDia-adjusted (Fig 2 spaghetti carries cohort_name; MAE overall)
  "dL_clean_Ipab_new", "dL_clean_sf2a_new", "dL_clean_sf3a_new",
  "dL_clean_sf6_new",  "dL_clean_sonnei_new",
  # case data, base/original times (observed points in Fig 4 / Fig 5)
  "dL_clean_Ipab", "dL_clean_sf2a", "dL_clean_sf3a", "dL_clean_sf6", "dL_clean_sonnei",
  # serotype case data (original times; MAE serospec + observed points)
  "dL_serotype_sf2a", "dL_serotype_sf3a", "dL_serotype_sonnei",
  # combined case data (MAE combined + Fig 5 panel D)
  "dL_combined_flexneri_sf2a_case", "dL_combined_flexneri_sf3a_case"
)

for (nm in main_rda) {
  load(file.path(path.expand(rda_dir), paste0(nm, ".rda")), envir = globalenv())
}

# raw-Excel availability flags (gate Table 1 + Figure 1 only)
have_compiled   <- file.exists(path.expand(raw_compiled))
have_table1_raw <- have_compiled &&
  file.exists(path.expand(raw_metadata)) && file.exists(path.expand(raw_durdia))
if (have_compiled) {
  compiled <- readxl::read_excel(path.expand(raw_compiled), sheet = sheet_compiled)
}
```

```{r}
#| label: shared-params-mae
#| message: false
#| warning: false

# Population kinetic-parameter summaries (== pop_table_sum / 2 / 3 in your .qmd)
sum_overall <- pop_param_summary(list(
  IpaB = overall_IpaB_pop_6, Sf2a = overall_Sf2a_pop_6, Sf3a = overall_Sf3a_pop_6,
  Sf6  = overall_Sf6_pop_6,  Sonnei = overall_Sonnei_pop_6))

sum_serospec <- pop_param_summary(list(
  Sf2a = serotype_sf2a_3, Sf3a = serotype_sf3a_3, Sonnei = serotype_sonnei_3))

sum_combined <- pop_param_summary(list(
  Sf2a = combined_flexneri_sf2a, Sf3a = combined_flexneri_sf3a))

# Per-individual MAE for each model class (drives Table 2 best-model pick + Table 4)
isos <- c("IgG", "IgA")

mae_overall <- dplyr::bind_rows(lapply(isos, function(io) dplyr::bind_rows(
  get_mae(overall_IpaB_pop_6,   dL_clean_Ipab_new,   "IpaB",   io),
  get_mae(overall_Sf2a_pop_6,   dL_clean_sf2a_new,   "Sf2a",   io),
  get_mae(overall_Sf3a_pop_6,   dL_clean_sf3a_new,   "Sf3a",   io),
  get_mae(overall_Sf6_pop_6,    dL_clean_sf6_new,    "Sf6",    io),
  get_mae(overall_Sonnei_pop_6, dL_clean_sonnei_new, "Sonnei", io)
)))

mae_serospec <- dplyr::bind_rows(lapply(isos, function(io) dplyr::bind_rows(
  get_mae(serotype_sf2a_3,   dL_serotype_sf2a,   "Sf2a",   io),
  get_mae(serotype_sf3a_3,   dL_serotype_sf3a,   "Sf3a",   io),
  get_mae(serotype_sonnei_3, dL_serotype_sonnei, "Sonnei", io)
)))

mae_combined <- dplyr::bind_rows(lapply(isos, function(io) dplyr::bind_rows(
  get_mae(combined_flexneri_sf2a, dL_combined_flexneri_sf2a_case, "Sf2a", io),
  get_mae(combined_flexneri_sf3a, dL_combined_flexneri_sf3a_case, "Sf3a", io)
)))
```

# Table 1 — Study population

> Needs the raw Excel (compiled + metadata + DurDia). Set the three paths in CONFIG to enable; otherwise this chunk is skipped.

```{r tbl1, eval=have_table1_raw, message=FALSE, warning=FALSE}
metadata <- readxl::read_excel(path.expand(raw_metadata), sheet = sheet_metadata)
durdia   <- readxl::read_excel(path.expand(raw_durdia),   sheet = sheet_durdia)

table1_study_population(build_table1_data(compiled, metadata, durdia))
```

# Figure 1 — Cross-reactivity heatmap

> Needs the raw compiled Excel (set `raw_compiled` in CONFIG); skipped otherwise.

```{r fig1, eval=have_compiled, fig.width=9, fig.height=6.5, message=FALSE, warning=FALSE}
figure_crossreactivity_heatmap(compiled)
```

# Figure 2 — Raw longitudinal IgG/IgA trajectories

```{r}
#| label: fig2-spaghetti
#| fig-width: 10
#| fig-height: 12
#| message: false
#| warning: false

# Uses the DurDia-adjusted objects because they carry `cohort_name`
# (infecting serotype) for the colour mapping.
figure_raw_spaghetti(
  case_list = list(
    IpaB   = dL_clean_Ipab_new,
    Sf2a   = dL_clean_sf2a_new,
    Sf3a   = dL_clean_sf3a_new,
    Sf6    = dL_clean_sf6_new,
    Sonnei = dL_clean_sonnei_new
  ),
  numeric_y = TRUE
)
```

# Table 2 — Kinetic parameters (best model per biomarker)

```{r}
#| label: tbl2
#| message: false
#| warning: false

best_lookup <- select_best_models(mae_overall, mae_serospec, mae_combined)
table2_kinetic_params(best_lookup, sum_overall, sum_serospec, sum_combined)
```

# Figure 4 — Individual-level model comparison

```{r}
#| label: fig4-comparison
#| fig-width: 10
#| fig-height: 13
#| message: false
#| warning: false

cmp_models <- list(
  overall_Sf2a    = overall_Sf2a_pop_6,
  serotype_sf2a   = serotype_sf2a_3,
  overall_Sonnei  = overall_Sonnei_pop_6,
  serotype_sonnei = serotype_sonnei_3,
  overall_Sf3a    = overall_Sf3a_pop_6,
  serotype_sf3a   = serotype_sf3a_3,
  combined_sf3a   = combined_flexneri_sf3a
)
cmp_datasets <- list(
  dL_clean_sf2a      = dL_clean_sf2a,
  dL_serotype_sf2a   = dL_serotype_sf2a,
  dL_clean_sonnei    = dL_clean_sonnei,
  dL_serotype_sonnei = dL_serotype_sonnei,
  dL_clean_sf3a      = dL_clean_sf3a,
  dL_serotype_sf3a   = dL_serotype_sf3a
)

# default sids = c("SOSAR-22008", "SOSAR-12005", "SOSAR-22007")
figure_individual_comparison(cmp_models, cmp_datasets)
```

# Table 3 — Age-stratified IpaB kinetic parameters

```{r}
#| label: tbl3
#| message: false
#| warning: false

table3_age_stratified(overall_IpaB_pop_under5, overall_IpaB_pop_plus5)
```

# Figure 5 — Population-level trajectories (panels A–E)

```{r}
#| label: fig5-trajectories
#| fig-width: 9
#| fig-height: 16
#| message: false
#| warning: false

traj_models <- list(
  IpaB            = overall_IpaB_pop_6,
  IpaB_under5     = overall_IpaB_pop_under5,
  IpaB_plus5      = overall_IpaB_pop_plus5,
  serotype_sonnei = serotype_sonnei_3,
  serotype_sf2a   = serotype_sf2a_3,
  combined_sf3a   = combined_flexneri_sf3a,
  overall_Sf6     = overall_Sf6_pop_6
)
traj_datasets <- list(
  dL_clean_Ipab      = dL_clean_Ipab,
  dL_serotype_sonnei = dL_serotype_sonnei,
  dL_serotype_sf2a   = dL_serotype_sf2a,
  dL_combined_sf3a   = dL_combined_flexneri_sf3a_case,
  dL_clean_sf6       = dL_clean_sf6
)

figure_population_trajectories(traj_models, traj_datasets)
```

# Table 4 — Per-individual MAE across modeling approaches

```{r}
#| label: tbl4
#| message: false
#| warning: false

t4 <- table4_mae(mae_overall, mae_serospec, mae_combined)
t4$flextable
```

```{r}
#| label: tbl4-footnote
#| echo: true

# The per-row "best for k/N" string (for the Results text / S-table):
cat(t4$winner_footnote, sep = "\n")
```
