---
title: "Chapter 1 — Supplement Figures & Tables"
subtitle: "Reproduces every supplement table & figure from library(shigella) + the regenerated .rda"
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    toc-location: right
editor: visual
---

```{=html}
<!--
PURPOSE
  GitHub / package-phase reproducibility document. With the shigella package
  installed, this renders every SUPPLEMENT table (S1–S4) and figure (S1–S6)
  straight from library(shigella) — no source() of R/.

HOW TO RUN
  1. Regenerate the fitted .rda once via data-raw/00..05 (script 05 writes the
     sensitivity_*.rda needed for the S3 table).
  2. Edit CONFIG: point `rda_dir` at that folder (it must also contain the
     sensitivity_*.rda).
  3. Render (Quarto) or run chunk-by-chunk. No raw Excel is needed here.
-->
```

```{r}
#| label: config
#| message: false

## ---- EDIT THIS -----------------------------------------------------------
rda_dir <- "~/Data/Manuscript" # *.rda objects + sensitivity_*.rda live here
## --------------------------------------------------------------------------
```

```{r}
#| label: setup
#| message: false
#| warning: false

# Everything comes from the installed package: shigella's Imports load with it,
# and the flextable/gt knit_print methods register on load. No source() of R/.
library(shigella)
```

```{r}
#| label: load-rda
#| message: false

if (!dir.exists(path.expand(rda_dir))) {
  cli::cli_abort(c(
    "{.var rda_dir} ({.path {rda_dir}}) does not exist.",
    "i" = "Set it to the folder containing the fitted .rda files in the CONFIG chunk."
  ))
}
supp_rda <- c(
  # overall + serotype + combined fits (sum_* , mae_* , S5 PPC, S6 forest, S3)
  "overall_IpaB_pop_6", "overall_Sf2a_pop_6", "overall_Sf3a_pop_6",
  "overall_Sf6_pop_6", "overall_Sonnei_pop_6",
  "serotype_sf2a_3", "serotype_sf3a_3", "serotype_sonnei_3",
  "combined_flexneri_sf2a", "combined_flexneri_sf3a",
  # case data, DurDia-adjusted (S5 PPC raw_dat + MAE overall + S3 overall data)
  "dL_clean_Ipab_new", "dL_clean_sf2a_new", "dL_clean_sf3a_new",
  "dL_clean_sf6_new", "dL_clean_sonnei_new",
  # case data, base/original times (S3 Fig observed points)
  "dL_clean_Ipab", "dL_clean_sf2a", "dL_clean_sf3a", "dL_clean_sf6",
  "dL_clean_sonnei",
  # serotype case data (MAE serospec + S3 table serotype data + S3 Fig)
  "dL_serotype_sf2a", "dL_serotype_sf3a", "dL_serotype_sonnei",
  # combined case data (MAE combined + S3 Fig panel C)
  "dL_combined_flexneri_sf2a_case", "dL_combined_flexneri_sf3a_case",
  # age-stratified case data, adjusted (S1 / S2 Fig spaghetti)
  "dL_clean_Ipab_new_under5", "dL_clean_sf2a_new_under5",
  "dL_clean_sf3a_new_under5",
  "dL_clean_sf6_new_under5", "dL_clean_sonnei_new_under5",
  "dL_clean_Ipab_new_plus5", "dL_clean_sf2a_new_plus5",
  "dL_clean_sf3a_new_plus5",
  "dL_clean_sf6_new_plus5", "dL_clean_sonnei_new_plus5"
)

for (nm in supp_rda) {
  load(file.path(path.expand(rda_dir), paste0(nm, ".rda")))
}
```

```{r}
#| label: shared-params-mae
#| message: false
#| warning: false

sum_overall <- pop_param_summary(list(
  IpaB = overall_IpaB_pop_6, Sf2a = overall_Sf2a_pop_6,
  Sf3a = overall_Sf3a_pop_6,
  Sf6 = overall_Sf6_pop_6, Sonnei = overall_Sonnei_pop_6
))
sum_serospec <- pop_param_summary(list(
  Sf2a = serotype_sf2a_3, Sf3a = serotype_sf3a_3, Sonnei = serotype_sonnei_3
))
sum_combined <- pop_param_summary(list(
  Sf2a = combined_flexneri_sf2a, Sf3a = combined_flexneri_sf3a
))

isos <- c("IgG", "IgA")
mae_overall <- dplyr::bind_rows(lapply(isos, function(io) {
  dplyr::bind_rows(
    get_mae(overall_IpaB_pop_6, dL_clean_Ipab_new, "IpaB", io),
    get_mae(overall_Sf2a_pop_6, dL_clean_sf2a_new, "Sf2a", io),
    get_mae(overall_Sf3a_pop_6, dL_clean_sf3a_new, "Sf3a", io),
    get_mae(overall_Sf6_pop_6, dL_clean_sf6_new, "Sf6", io),
    get_mae(overall_Sonnei_pop_6, dL_clean_sonnei_new, "Sonnei", io)
  )
}))
mae_serospec <- dplyr::bind_rows(lapply(isos, function(io) {
  dplyr::bind_rows(
    get_mae(serotype_sf2a_3, dL_serotype_sf2a, "Sf2a", io),
    get_mae(serotype_sf3a_3, dL_serotype_sf3a, "Sf3a", io),
    get_mae(serotype_sonnei_3, dL_serotype_sonnei, "Sonnei", io)
  )
}))
mae_combined <- dplyr::bind_rows(lapply(isos, function(io) {
  dplyr::bind_rows(
    get_mae(combined_flexneri_sf2a, dL_combined_flexneri_sf2a_case, "Sf2a", io),
    get_mae(combined_flexneri_sf3a, dL_combined_flexneri_sf3a_case, "Sf3a", io)
  )
}))
```

# S1 Table — Prior distributions

```{r}
#| label: s1-table
table_s1_priors()
```

# S2 Table — MCMC settings

```{r}
#| label: s2-table
table_s2_mcmc()
```

# S3 Table — Prior-specification robustness (MAE)

> Reads `sensitivity_*.rda` directly from `rda_dir` and recomputes per-model MAE.

```{r}
#| label: s3-table
#| message: false
#| warning: false

s3_datasets <- list(
  dL_clean_sonnei_new = dL_clean_sonnei_new,
  dL_clean_sf3a_new   = dL_clean_sf3a_new,
  dL_clean_sf2a_new   = dL_clean_sf2a_new,
  dL_serotype_sonnei  = dL_serotype_sonnei,
  dL_serotype_sf3a    = dL_serotype_sf3a,
  dL_serotype_sf2a    = dL_serotype_sf2a
)

build_table_s3(dir = rda_dir, datasets = s3_datasets)
```

# S4 Table — All models (overall + serotype + combined)

```{r}
#| label: s4-table
#| message: false
#| warning: false

table_s4_all_models(
  sum_overall, sum_serospec, sum_combined,
  models_overall = list(
    IpaB   = overall_IpaB_pop_6,  Sf2a = overall_Sf2a_pop_6,
    Sf3a   = overall_Sf3a_pop_6,  Sf6  = overall_Sf6_pop_6,
    Sonnei = overall_Sonnei_pop_6
  ),
  models_serospec = list(
    Sf2a = serotype_sf2a_3, Sf3a = serotype_sf3a_3, Sonnei = serotype_sonnei_3
  ),
  models_combined = list(
    Sf2a = combined_flexneri_sf2a, Sf3a = combined_flexneri_sf3a
  )
)
```

# S1 Fig — Age-stratified raw trajectories (\< 5 years)

```{r}
#| label: s1-fig
#| fig-width: 10
#| fig-height: 12
#| message: false
#| warning: false

figure_s1_age_spaghetti(list(
  IpaB   = dL_clean_Ipab_new_under5,
  Sf2a   = dL_clean_sf2a_new_under5,
  Sf3a   = dL_clean_sf3a_new_under5,
  Sf6    = dL_clean_sf6_new_under5,
  Sonnei = dL_clean_sonnei_new_under5
))
```

# S2 Fig — Age-stratified raw trajectories (\>= 5 years)

```{r}
#| label: s2-fig
#| fig-width: 10
#| fig-height: 12
#| message: false
#| warning: false

figure_s2_age_spaghetti(list(
  IpaB   = dL_clean_Ipab_new_plus5,
  Sf2a   = dL_clean_sf2a_new_plus5,
  Sf3a   = dL_clean_sf3a_new_plus5,
  Sf6    = dL_clean_sf6_new_plus5,
  Sonnei = dL_clean_sonnei_new_plus5
))
```

# S3 Fig — Additional individual comparisons

```{r}
#| label: s3-fig
#| fig-width: 10
#| fig-height: 13
#| message: false
#| warning: false

cmp_models <- list(
  overall_Sf2a    = overall_Sf2a_pop_6,
  serotype_sf2a   = serotype_sf2a_3,
  overall_Sonnei  = overall_Sonnei_pop_6,
  serotype_sonnei = serotype_sonnei_3,
  overall_Sf3a    = overall_Sf3a_pop_6,
  serotype_sf3a   = serotype_sf3a_3,
  combined_sf3a   = combined_flexneri_sf3a
)
cmp_datasets <- list(
  dL_clean_sf2a      = dL_clean_sf2a,
  dL_serotype_sf2a   = dL_serotype_sf2a,
  dL_clean_sonnei    = dL_clean_sonnei,
  dL_serotype_sonnei = dL_serotype_sonnei,
  dL_clean_sf3a      = dL_clean_sf3a,
  dL_serotype_sf3a   = dL_serotype_sf3a
)

# default sids = c("SOSAR-23004", "SOSAR-21018", "SOSAR-21002")
figure_s3_additional_comparison(cmp_models, cmp_datasets)
```

# S4 Fig — Per-individual MAE slopegraph

```{r}
#| label: s4-fig
#| fig-width: 10
#| fig-height: 10
#| message: false
#| warning: false

figure_s4_mae_slopegraph(mae_overall, mae_serospec, mae_combined)
```

# S5 Fig — Posterior predictive checks

```{r}
#| label: s5-ipab
#| fig-width: 6.83
#| fig-height: 4.5
#| message: false
#| warning: false
posterior_pred(overall_IpaB_pop_6, dL_clean_Ipab_new,
  by_antigen = TRUE,
  n_sim = 25, antigen_name = "IpaB"
)
```

```{r}
#| label: s5-sf2a
#| fig-width: 6.83
#| fig-height: 4.5
#| message: false
#| warning: false
posterior_pred(overall_Sf2a_pop_6, dL_clean_sf2a_new,
  by_antigen = TRUE,
  n_sim = 25, antigen_name = "Sf2a"
)
```

```{r}
#| label: s5-sf3a
#| fig-width: 6.83
#| fig-height: 4.5
#| message: false
#| warning: false
posterior_pred(overall_Sf3a_pop_6, dL_clean_sf3a_new,
  by_antigen = TRUE,
  n_sim = 25, antigen_name = "Sf3a"
)
```

```{r}
#| label: s5-sf6
#| fig-width: 6.83
#| fig-height: 4.5
#| message: false
#| warning: false
posterior_pred(overall_Sf6_pop_6, dL_clean_sf6_new,
  by_antigen = TRUE,
  n_sim = 25, antigen_name = "Sf6"
)
```

```{r}
#| label: s5-sonnei
#| fig-width: 6.83
#| fig-height: 4.5
#| message: false
#| warning: false
posterior_pred(overall_Sonnei_pop_6, dL_clean_sonnei_new,
  by_antigen = TRUE,
  n_sim = 25, antigen_name = "Sonnei"
)
```

# S6 Fig — Kinetic-parameter forest (model comparison)

```{r}
#| label: s6-fig
#| fig-width: 10
#| fig-height: 15
#| message: false
#| warning: false

plot_model_comparison_forest(
  pop_overall  = sum_overall,
  pop_serospec = sum_serospec,
  pop_combined = sum_combined
)
```
