---
title: "Chapter 2 — Correlated Multivariate Antibody Kinetics"
subtitle: "A Kronecker-structured Bayesian hierarchical model: methodology and Phase 0/1 simulation diagnostics"
author: "Kwan Ho Lee"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    theme: cosmo
    code-fold: true
    fig-width: 7
    fig-height: 4.5
  docx:
    
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

```{r setup}
#| include: false
options(knitr.kable.NA = "")  # nolint: undesirable_function_linter
```

# Introduction {#sec-intro}

## Motivation

Chapter 1 established a univariate Bayesian hierarchical model for the
post-symptom-onset antibody trajectory of a single antigen-isotype
biomarker, fitting each biomarker independently. This independence
assumption is biologically unrealistic: within the same subject, IgG and
IgA responses share the same infection event, the same immune system,
and the same antigenic stimulation, so their kinetic parameters are
expected to co-vary. Ignoring this dependence discards information that
could improve parameter estimation precision and downstream
seroincidence inference.

## Aim of Chapter 2

Chapter 2 extends the Chapter 1 framework to a multivariate hierarchical
model in which the random-effect covariance of the per-subject log-scale
kinetic parameters has a Kronecker product structure,
$\Sigma = \Sigma_B \otimes \Sigma_P$. Here $\Sigma_B$ captures
correlation between biomarkers and $\Sigma_P$ captures correlation
between kinetic parameters, reducing the unconstrained $5K \times 5K$
joint covariance to a structured, interpretable factorization.

## Scope of this document

@sec-methods derives the model and maps each mathematical object to its
implementation in the `shigella` R package and the Stan model file.
@sec-results presents the Phase 0 (interactive SLURM) and Phase 1 (batch
SLURM) reproducibility check at two sample sizes relevant to the Chapter
1 cohort. @sec-discussion interprets the n = 48 fit pathology and
proposes next steps.

# Methods {#sec-methods}

## The Chapter 1 univariate kinetic model

Following Teunis et al. (2016), the post-symptom-onset antibody
concentration $y(t)$ in a subject's blood at time $t$ is modeled as a
deterministic two-phase trajectory. This formulation builds on the
within-host rise--decay seroresponse models developed by de Graaf et al.
(2014) and Teunis et al. (2016):

$$
y(t) \;=\;
\begin{cases}
y_0 \exp(\beta t), & 0 \le t \le t_1 \quad \text{(rise phase)} \\[6pt]
\left[ y_1^{1-r} - (1-r)\,\alpha\,(t - t_1) \right]^{1/(1-r)}, & t > t_1 \quad \text{(decay phase)}
\end{cases}
$$

where $\beta = \log(y_1 / y_0) / t_1$. The five per-subject kinetic
parameters are $(y_0, y_1, t_1, \alpha, r)$, denoting baseline level,
peak level, time-to-peak, decay rate, and decay shape.

For hierarchical modeling, these are reparameterized into unconstrained
log-scale coordinates:

$$
\boldsymbol{\theta}_{i,k}
\;=\;
\bigl( \log y_{0}, \;\; \log(y_1 - y_0), \;\; \log t_1, \;\; \log\alpha, \;\; \log(r - 1) \bigr)_{i,k}
\;\in\; \mathbb{R}^{5},
$$

for subject $i$ and biomarker $k$.

## The multivariate extension {#sec-mv}

Let $K$ index biomarkers (antigen-isotype pairs; $K=2$ for the IgG/IgA
pair) and $P = 5$ the number of kinetic parameters. Stack the
per-subject log-scale parameters across biomarkers:

$$
\boldsymbol{\theta}_i \;=\;
\bigl( \boldsymbol{\theta}_{i,1}^{\top},\,\boldsymbol{\theta}_{i,2}^{\top},\,\ldots,\,\boldsymbol{\theta}_{i,K}^{\top} \bigr)^{\top}
\;\in\; \mathbb{R}^{KP}.
$$

The Chapter 2 random-effect prior is

$$
\boldsymbol{\theta}_i \;\sim\; \mathcal{N}_{KP}\bigl( \boldsymbol{\mu},\; \Sigma_B \otimes \Sigma_P \bigr),
\qquad i = 1, \ldots, n,
$$ {#eq-kron}

where $\otimes$ denotes the Kronecker product,
$\Sigma_B \in \mathbb{R}^{K \times K}$ is the between-biomarker
covariance, and $\Sigma_P \in \mathbb{R}^{P \times P}$ is the
between-parameter covariance. This is the multivariate hierarchical
parameterization used to separate covariance across response dimensions
from covariance across kinetic parameters (Gelman et al., 2014).

The Kronecker structure factorizes the cross-(biomarker, parameter)
covariance as

$$
\mathrm{Cov}(\theta_{i, k, p},\, \theta_{i, k', p'})
\;=\; \Sigma_B[k, k'] \cdot \Sigma_P[p, p'].
$$

## Identifiability and the LKJ–scale parameterization

The decomposition $\Sigma_B \otimes \Sigma_P$ is non-unique up to a
scalar: for any $c \neq 0$, the pair $(c\Sigma_B, c^{-1}\Sigma_P)$
yields the same product. To resolve this, $\Sigma_B$ and $\Sigma_P$ are
constrained to be correlation matrices and the marginal variances
are absorbed into separate positive scale vectors:

$$
\Sigma_B \;=\; \mathrm{diag}(\boldsymbol{\tau}_B)\,\Omega_B\,\mathrm{diag}(\boldsymbol{\tau}_B),
\quad
\Sigma_P \;=\; \mathrm{diag}(\boldsymbol{\tau}_P)\,\Omega_P\,\mathrm{diag}(\boldsymbol{\tau}_P),
$$

with $\Omega_B, \Omega_P$ correlation matrices (unit diagonal) and
$\boldsymbol{\tau}_B \in \mathbb{R}_+^K$,
$\boldsymbol{\tau}_P \in \mathbb{R}_+^P$.

## Priors

| Parameter | Prior | Rationale |
|---------------------------|------------------|---------------------------|
| $\boldsymbol{\mu} \in \mathbb{R}^{KP}$ | $\mathcal{N}(\mathbf{0}, 5\,\mathbf{I})$ componentwise | Weakly informative, on the log-scale parameter space |
| $\Omega_B$ | $\mathrm{LKJ}(\eta = 2)$ | Mild concentration around independence; permits either sign of correlation |
| $\Omega_P$ | $\mathrm{LKJ}(\eta = 2)$ | Same rationale |
| $\boldsymbol{\tau}_B$ | $\mathrm{half\text{-}}\mathcal{N}(0,\,1)$ componentwise | Positive scale; weakly informative |
| $\boldsymbol{\tau}_P$ | $\mathrm{half\text{-}}\mathcal{N}(0,\,1)$ componentwise | Same |

The LKJ prior is used because it defines a proper prior over correlation
matrices and allows transparent control over concentration around the
identity matrix (Lewandowski et al., 2009). I do not impose a sign
constraint on $\rho_B$: although IgG and IgA responses share the same
infection event, class switching and mucosal/systemic response timing
can differ, so early post-infection data may plausibly support either
positive or negative within-subject IgG--IgA correlation (Stavnezer et
al., 2008; Mattoo & Cherry, 2005).

The target estimand of interest for this calibration study is the
off-diagonal entry $\rho_B \equiv \Omega_B[1,2]$, the between-biomarker
correlation.

## Mapping mathematics to code {#sec-code-map}

| Mathematical object | Implemented in |
|------------------------------------|------------------------------------|
| Two-phase trajectory $y(t)$ + log-scale parameterization | `inst/stan/model_2.stan` (transformed parameters block) |
| Synthetic data drawn from @eq-kron with known $\rho_B$ | `R/sim_correlated_case_data.R` |
| Construction of $\Sigma_B \otimes \Sigma_P$ from $(\Omega_B, \Omega_P, \boldsymbol{\tau}_B, \boldsymbol{\tau}_P)$ | `R/build_sigma_matrices.R` |
| Per-subject parameter draws $\boldsymbol{\theta}_i$ | `R/draw_subject_params.R` |
| Mean trajectory $\log \mu_k(t)$ | `R/compute_log_mu_k.R` |
| Stan data list assembly | `R/prep_data_stan.R` |
| Prior hyperparameter list | `R/prep_priors_stan.R` |
| MCMC fit invocation (cmdstanr) | `R/run_mod_stan.R` |
| Posterior summary + diagnostics | `R/postprocess_stan_output.R`, `R/summarize_matrix_array.R` |
| Phase 0 (interactive SLURM) workflow | `R/run_phase0_diagnostic.R` |
| Phase 1 (batch SLURM) workflow | `R/run_phase1_diagnostic.R` |

The Stan implementation should use a non-centered hierarchical
parameterization, because non-centered forms often reduce difficult
funnel-like posterior geometry in hierarchical models fitted with HMC
(Betancourt & Girolami, 2015). In code, this means sampling
standard-normal latent variables and transforming them through the
Cholesky factor of the structured covariance matrix rather than sampling
subject-level parameters directly.

## Simulation design

The package function `sim_correlated_case_data()` generates data from
@eq-kron with a user-specified ground-truth $\rho_B$. Two sample sizes
are tested here:

| Setting | $n$ subjects | Per-subject obs. | $\rho_B$ truth | Warmup / Sampling per chain | Chains | Seed |
|-----------|----------:|----------:|----------:|----------:|----------:|----------:|
| Light | 5 | 10 | $+0.6$ | 500 / 500 | 2 | 20260513 |
| Realistic | 48 | 10 | $+0.6$ | 1000 / 1000 | 2 | 20260513 |

The two sample sizes are chosen as a sanity check ($n=5$) and as the
Chapter-1 IpaB analytical cohort size ($n=48$). Each fit is run twice
with identical seed and model file but in two different execution modes
(interactive `salloc` vs batch `sbatch`) to verify execution determinism
(see @sec-results).

For the later multi-replicate simulation study, performance summaries
should be reported with Monte Carlo standard errors, rather than
interpreted from a small pilot alone. Following Morris et al. (2019),
the planned full simulation should summarize bias, mean squared error,
coverage, and the Monte Carlo standard error of each performance
measure.

# Results {#sec-results}

Sampler diagnostics are interpreted using standard HMC/NUTS guidance:
divergences indicate numerical integration failures in difficult
posterior geometry, maximum-treedepth saturation indicates that NUTS
repeatedly required the maximum allowed trajectory length, and
$\widehat R$ plus effective sample size summarize cross-chain
convergence and Monte Carlo precision (Hoffman & Gelman, 2014;
Betancourt, 2018; Vehtari et al., 2021).

## Reproducibility at $n = 5$ {#sec-results-n5}

@tbl-n5 reports the side-by-side comparison of Phase 0 and Phase 1 at
$n = 5$.

```{r}
#| label: tbl-n5
#| tbl-cap: "Phase 0 (interactive `salloc`) vs Phase 1 (batch `sbatch`) at 
#| $n = 5$, $\\rho_B^{\\text{truth}} = +0.6$. Both phases use identical seed 
#| and model file."

cmp_n5 <- data.frame(
  Metric         = c("Status", "Elapsed (min)", "Posterior median $\\rho_B$",
                     "95% CrI lower (2.5%)", "95% CrI upper (97.5%)",
                     "ESS_bulk", "$\\hat R$", "Divergent transitions (/1000)",
                     "Max-treedepth hits (/1000)"),
  Phase0 = c("OK", "9.54", "$-0.034$", "$-0.463$", "$+0.611$",
             "6", "1.371", "13", "493"),
  Phase1 = c("OK", "9.40", "$-0.034$", "$-0.463$", "$+0.611$",
             "6", "1.371", "13", "493")
)

knitr::kable(cmp_n5, 
             col.names = c("Metric", "Phase 0 (interactive)", 
                           "Phase 1 (batch)"), 
             align = "lrr")
```

All numerical summaries are bit-for-bit identical between the two
execution modes. Stan's HMC sampler is deterministic given identical
seed, model, data, and runtime, so this match confirms that:

1.  The cluster environment (SLURM batch vs interactive allocation,
    compute-node `/tmp`, library paths) does not affect the fit.
2.  Any pathology observed at higher $n$ is therefore a property of the
    model–data combination, not an environment artifact.

The $n = 5$ fit itself is uninformative ($\hat R = 1.37$, ESS = 6): at
this sample size the LKJ(2) prior dominates the posterior, which is
expected. The $n = 5$ run serves only as a reproducibility sanity check.

## Reproducibility and fit pathology at $n = 48$ {#sec-results-n48}

@tbl-n48 reports the same comparison at $n = 48$, the Chapter 1 IpaB
cohort size.

```{r}
#| label: tbl-n48
#| tbl-cap: "Phase 0 (interactive `salloc`) vs Phase 1 (batch `sbatch`) at 
#| $n = 48$, $\\rho_B^{\\text{truth}} = +0.6$. Identical seed and model. Cells 
#| flagged violate conventional diagnostic thresholds."

cmp_n48 <- data.frame(
  Metric = c("Status", "Elapsed (min)", "Posterior median $\\rho_B$",
             "95% CrI lower (2.5%)", "95% CrI upper (97.5%)",
             "ESS_bulk", "$\\hat R$", "Divergent transitions (/2000)",
             "Max-treedepth hits (/2000)"),
  Phase0 = c("OK", "144.54", "$-0.624$", "$-0.884$", "$-0.378$ ",
             "2", "2.824 ", "1", "1999 "),
  Phase1 = c("OK", "131.83", "$-0.624$", "$-0.884$", "$-0.378$",
             "2", "2.824 ", "1", "1999 "),
  Acceptable = c("OK", "—", "near $+0.6$", "—", "should contain $+0.6$",
                 "$> 400$", "$< 1.01$", "$< 5\\%$", "$< 10\\%$")
)
knitr::kable(cmp_n48,
             col.names = c("Metric", "Phase 0 (interactive)",
                           "Phase 1 (batch)", "Acceptable"),
             align = "lrrr")
```

Three observations from @tbl-n48 are central to the interpretation in
@sec-discussion:

1.  Phase 0 ≡ Phase 1 at$n = 48$ as well: all summaries match
    bit-for-bit. Environment is again ruled out.
2.  Posterior has the wrong sign: the true value is $\rho_B = +0.6$ but
    the recovered median is $-0.624$. The 95% credible interval
    $[-0.884,\,-0.378]$ does not contain the truth.
3.  Severe sampler geometry pathology, but few divergences:
    $\hat R = 2.82$ and 99.95% of transitions hit the maximum tree
    depth, yet the divergent count is only 1 / 2000 (0.05%). This is a
    specific signature: the sampler is unable to traverse the typical
    set within the step budget, rather than falling into a funnel.

# Discussion {#sec-discussion}

## Departure from the over-parameterization pattern

A naive prior expectation was that $n = 48$ would show weak
identification — the canonical symptom being a posterior that is wide
and roughly centered on the LKJ prior, with $\hat R \approx 1$ (chains
agree they are uninformative) and credible-interval coverage at the
nominal level.

The observed pattern is qualitatively different:

| Feature | Standard weak identification | Observed at $n = 48$ |
|------------------------|------------------------|------------------------|
| Posterior shape | Wide, prior-like | Narrow (sd ≈ 0.24), confidently wrong |
| 95% CrI for truth | Wide, covers truth | Narrow, excludes truth |
| Between-chain agreement | $\hat R \approx 1$ | $\hat R = 2.82$ |
| Treedepth hits | Moderate | 99.95% |
| Divergent count | Nonzero | 0.05% |

The combination near-zero divergent count + near-100% treedepth +
high$\hat R$ points away from "data uninformative for $\rho_B$" and
toward "sampler trapped in a low-information region of parameter space."

## Three candidate explanations

After ruling out environment and code-determinism effects via
@sec-results, three candidate mechanisms remain.

### Structural identifiability of $\Sigma_B$ vs $\Sigma_P$ at finite $n$

The Kronecker decomposition is asymptotically identified up to the
scalar ambiguity resolved by the correlation-matrix constraint
(@sec-mv). At infinite $n$ the data separate $\Omega_B$ from $\Omega_P$.
At finite $n$, however, multiple $(\Omega_B, \Omega_P)$ pairs may yield
observationally near-equivalent products $\Omega_B \otimes \Omega_P$.
The boundary of empirical identifiability under our priors and
observation density at $K = 2, P = 5$ may exceed $n = 48$.

### Likelihood multimodality

The LKJ(2) prior on a $K = 2$ correlation matrix is unimodal in
$\rho_B$, but the likelihood for $\Omega_B$ given a finite dataset may
not be. With only two chains, a mode-finding sampler could locate two
distinct modes that yield large $\hat R$ even after warmup, as appears
to be the case here.

### Simulation–model mismatch

The least glamorous and most consequential possibility: that
`sim_correlated_case_data()` does not actually produce data drawn from
the exact structure that `inst/stan/model_2.stan` assumes. Candidate
sources of mismatch include Cholesky-factor ordering conventions,
biomarker-vs-parameter axis ordering in the Kronecker product, or
noise-model parameterization. This must be checked first, because if
confirmed it would make the other two analyses moot.

## Open questions

Distinguishing among the three mechanisms above will require further
investigation, and the appropriate sequencing of those investigations is
itself a question on which advisor input is welcomed. A natural next
step is a targeted simulation-based calibration workflow: simulate from
known parameter values, refit the model, and verify that posterior
summaries recover the known truth across repeated datasets (Talts et
al., 2018).

## Scope notes

-   The findings here pertain to Model 2 at $n = 48$ under the specific
    prior and simulation configuration described in @sec-methods.
    Implications for the Chapter 1 univariate model, if any, are outside
    the scope of this document.
-   Only one true value of $\rho_B$ ($+0.6$) and only $K = 2$ biomarkers
    were tested. Phase 0 and Phase 1 use the same seed.

# Appendix A — Software environment {.appendix .unnumbered}

| Component           | Version                                     |
|---------------------|---------------------------------------------|
| R                   | 4.6.0 (2026-04-24)                          |
| cmdstanr            | 0.8.0                                       |
| cmdstan             | 2.38.0                                      |
| posterior           | 1.7.0                                       |
| shigella            | 0.0.0.9006                                  |
| serodynamics        | 0.0.0.9050                                  |
| Compute environment | UC Davis Shiva HPC (`r_chapter2` conda env) |

# Appendix B — Reproducibility {.appendix .unnumbered}

Branch `chapter2-stan-simulation` of
`https://github.com/UCD-SERG/shigella`.

``` bash
# Phase 0 (interactive SLURM via salloc)
salloc --time=04:00:00 --cpus-per-task=2 --mem=20G
Rscript scripts/phase0_interactive_reproducibility.R       # n=5
Rscript scripts/phase0_interactive_reproducibility_n48.R   # n=48
exit

# Phase 1 (batch SLURM)
sbatch --time=04:00:00 --cpus-per-task=2 --mem=20G \
       --output=logs/phase1/phase1_%j.out \
       slurm/phase1_single.sbatch         # n=5
sbatch --time=08:00:00 --cpus-per-task=2 --mem=20G \
       --output=logs/phase1/phase1_n48_%j.out \
       slurm/phase1_single_n48.sbatch     # n=48
```

Result files: `outputs/phase{0,1}/one_fit_n{5,48}*.rds` and comparison
tables `outputs/phase1/p0_vs_p1_comparison_<JOBID>.rds`.

# References {.unnumbered}

::: {#refs}
Betancourt, M. (2018). *A conceptual introduction to Hamiltonian Monte
Carlo*. arXiv:1701.02434. https://doi.org/10.48550/arXiv.1701.02434

Betancourt, M., & Girolami, M. (2015). Hamiltonian Monte Carlo for
hierarchical models. In *Current Trends in Bayesian Methodology with
Applications* (pp. 79--101). Chapman & Hall/CRC.

de Graaf, W. F., Kretzschmar, M. E. E., Teunis, P. F. M., & Diekmann, O.
(2014). *A two-phase within-host model for immune response and its
application to serological profiles of pertussis*. Epidemics, 9, 1--7.
https://doi.org/10.1016/j.epidem.2014.08.002

Gelman, A., Carlin, J. B., Stern, H. S., Dunson, D. B., Vehtari, A., &
Rubin, D. B. (2014). *Bayesian Data Analysis* (3rd ed.). Chapman &
Hall/CRC.

Hoffman, M. D., & Gelman, A. (2014). *The No-U-Turn Sampler: Adaptively
setting path lengths in Hamiltonian Monte Carlo*. Journal of Machine
Learning Research, 15, 1593--1623.

Lewandowski, D., Kurowicka, D., & Joe, H. (2009). *Generating random
correlation matrices based on vines and extended onion method*. Journal
of Multivariate Analysis, 100(9), 1989--2001.
https://doi.org/10.1016/j.jmva.2009.04.008

Mattoo, S., & Cherry, J. D. (2005). *Molecular pathogenesis,
epidemiology, and clinical manifestations of respiratory infections due
to Bordetella pertussis and other Bordetella subspecies*. Clinical
Microbiology Reviews, 18(2), 326--382.
https://doi.org/10.1128/CMR.18.2.326-382.2005

Morris, T. P., White, I. R., & Crowther, M. J. (2019). *Using simulation
studies to evaluate statistical methods*. Statistics in Medicine,
38(11), 2074--2102. https://doi.org/10.1002/sim.8086

Stavnezer, J., Guikema, J. E. J., & Schrader, C. E. (2008). *Mechanism
and regulation of class switch recombination*. Annual Review of
Immunology, 26, 261--292.
https://doi.org/10.1146/annurev.immunol.26.021607.090248

Talts, S., Betancourt, M., Simpson, D., Vehtari, A., & Gelman, A.
(2018). *Validating Bayesian inference algorithms with simulation-based
calibration*. arXiv:1804.06788.
https://doi.org/10.48550/arXiv.1804.06788

Teunis, P. F. M., van Eijkeren, J. C. H., de Graaf, W. F., Bonačić
Marinović, A., & Kretzschmar, M. E. E. (2016). *Linking the seroresponse
to infection to within-host heterogeneity in antibody production*.
Epidemics, 16, 33--39. https://doi.org/10.1016/j.epidem.2016.04.001

Vehtari, A., Gelman, A., Simpson, D., Carpenter, B., & Bürkner, P.-C.
(2021). *Rank-normalization, folding, and localization: An improved*
$\widehat R$ for assessing convergence of MCMC. Bayesian Analysis,
16(2), 667--718. https://doi.org/10.1214/20-BA1221
:::
