---
title: "Priors Analysis of Shigella"
author: "Kwan Ho Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    toc_float: true
    #number_sections: true  ## if you want number sections at each table header
    #theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r echo=F, include=F, warning=F, notes = F}

library(dplyr)
library(runjags)
library(readr)
library(serodynamics)
library(serocalculator)
library(tidyverse)
library(runjags)
library(coda)
library(ggmcmc)
library(ggbeeswarm)
library(ggpubr)
library(knitr)
library(table1)
library(readxl)
devtools::load_all()

options(scipen = 99, digits = 4)

#model file
file.mod <- fs::path_package("serodynamics", "extdata/model.jags")
```

# Load the processed Shigella dataset

```{r}
#| label: "load shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'
df <- 
  fs::path_package(
    package = "shigella",
    "extdata/3.8.2024 Compiled Shigella datav2.xlsx"
  ) |> 
  read_excel(sheet = "Compiled")
```

## Focus on Ipab_IgG

```{r}
#| label: "manipulate Shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

# filter ipab
dL_clean_Ipab <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_ipab_MFI
)

dL_clean_Ipab_IgG<-dL_clean_Ipab %>%
  filter(antigen_iso == "IgG")

```

# Analaysis overview:

In this analysis, we are examining how **changing hyperpriors** for

-   `prec_hyp_param` (spread/uncertainty) and

-   `mu_hyp_param` (center/average guess)

affects **R̂ values (convergence)** and **parameter estimates** in our Bayesian hierarchical model.

The Shigella dataset is filtered to include only the **IpaB_IgG biomarker**. All JAGS models were run with the following constants:

-   `nchain = 4`, `nadapt = 1000`, `nburn = 500`, `nmc = 2500`, `niter = 5000`.

# Test 2: Sensitivity analysis of prior mean (`mu_hyp_param`)

**Goal**: After choosing reasonable spreads from Test 1, check how shifting the **center** of the priors affects convergence and estimates. This tests robustness of the model to different plausible prior centers.

**Setup**:

-   Use the optimized precision settings from Test 1.

-   Sweep each `mu_hyp_param` individually over a plausible grid (e.g., $y_0 = \{2.5, 3.0, 3.5, 4.0, 4.5\}$).

**Method**:

-   Vary the mean prior for one parameter at a time, holding the others fixed.

-   For each run, check $\hat{R}$ and parameter estimates.

**Interpretation**:

-   If results remain stable as the mean is shifted, the data dominate (robust to prior).

-   If results change dramatically, the prior center matters and should be chosen carefully.

