---
title: "Priors Analysis of Shigella2"
subtitle: "Test 1: Sensitivity Analysis of Prior Precision for Shigella (prec_hyp_param)"
author: "Kwan Ho Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true             # enable table of contents
    toc_depth: 3          # up to three heading levels (#, ##, ###)
    toc_float: true       # floating TOC on the side
    highlight: tango      # syntax highlighting style
  pdf_document:
    toc: true             # also enable table of contents for PDF
    toc_depth: 3          # same heading depth for PDF
    highlight: tango      # consistent highlighting
---

```{r echo=F, include=F, warning=F, notes = F}

library(dplyr)
library(runjags)
library(readr)
library(serodynamics)
library(serocalculator)
library(tidyverse)
library(runjags)
library(coda)
library(ggmcmc)
library(ggbeeswarm)
library(ggpubr)
library(knitr)
library(table1)
library(readxl)
devtools::load_all()

options(scipen = 99, digits = 4)

#model file
file.mod <- fs::path_package("serodynamics", "extdata/model.jags")
```

```{r}
#| label: "load shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'
df <- 
  fs::path_package(
    package = "shigella",
    "extdata/3.8.2024 Compiled Shigella datav2.xlsx"
  ) |> 
  read_excel(sheet = "Compiled")
```

```{r}
#| label: "manipulate Shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

# filter ipab
dL_clean_Ipab <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_ipab_MFI
)

dL_clean_Ipab_IgG<-dL_clean_Ipab %>%
  filter(antigen_iso == "IgG")

```

# Analaysis overview:

In this analysis, we are examining how **changing hyperpriors** for

-   `prec_hyp_param` (spread/uncertainty) and

-   `mu_hyp_param` (center/average guess)

affects **R̂ values (convergence)** and **parameter estimates** in our Bayesian hierarchical model.

The Shigella dataset is filtered to include only the **IpaB_IgG biomarker**. All JAGS models were run with the following constants:

-   `nchain = 4`, `nadapt = 1000`, `nburn = 500`, `nmc = 2500`, `niter = 5000`.

# Test 1: Sensitivity analysis of prior spread (`prec_hyp_param`)

**Goal**: Check how the model behaves when we change the **prior precision** (spread) for each parameter, one at a time. This tests whether the model is sensitive to overly tight or overly vague priors, and helps identify a safe level of spread that supports good convergence.

**Setup**:

-   Default priors for means:

    `mu.hyp = c(1.0, 7.0, 1.0, -2.0, -3.0)`

-   Precision sweep values:

    `prec.hyp = {10, 5, 2, 1, 0.1, 0.01, 0.0001, 0.000001}`

-   Other hyperpriors:

    `omega = diag(c(10,10,10,10,10,10)), wishdf = 20, prec.logy.hyp = c(4.0,1)`

**Method**:

-   Vary the precision for one parameter at a time (e.g., $y_0$), while holding the other four precisions fixed.

-   For each run, record how $\hat{R}$ and parameter estimates for all parameters respond.

**Interpretation**:

-   High precision (narrow prior) forces estimates close to the mean guess.

-   Low precision (diffuse prior) gives more freedom but may cause poor convergence.

-   The goal is to identify a stable range of precision values that allow good convergence ($\hat{R} \approx 1$, stable ESS) without over-constraining the model.

------------------------------------------------------------------------


```{r echo=F, warning=F, notes=F, include = F}
# --- summarize with serodynamics::post_summ() by prior value -----------------
post_summ_by_prior <- function(data,
                               iso   = unique(data$Iso_type),
                               param = unique(data$Parameter),
                               strat = unique(data$Stratification)) {
  # runs post_summ() separately for each Priors value and binds results
  data %>%
    dplyr::group_by(Priors) %>%
    dplyr::group_modify(~ serodynamics::post_summ(.x, iso = iso,
                                                  param = param, strat = strat)) %>%
    dplyr::ungroup()
}

mu_vec <- c(y0 = 1.0, y1 = 7.0, t1 = 1.0, alpha = -2.0, shape = -3.0)  # from setup
```



```{r echo=F, warning=F, notes=F, include = F}
# ---------- helper: 95% prior ribbon on the raw plotting scale ----------
.make_prior_ribbon <- function(df_params_priors, mu_hyp, mu_scale = c("log","log10","identity")[1],
                               p_min_for_ribbon = NULL) {
  if (is.null(mu_hyp)) return(NULL)
  mu_scale <- match.arg(mu_scale)

  # df_params_priors must have: Parameter, Priors (precision grid)
  keep <- intersect(df_params_priors$Parameter, names(mu_hyp))
  if (length(keep) == 0) return(NULL)

  df <- dplyr::filter(df_params_priors, .data$Parameter %in% keep)

  # optional clipping of tiny precisions (helps avoid absurd ranges)
  if (!is.null(p_min_for_ribbon)) {
    df <- dplyr::filter(df, .data$Priors >= p_min_for_ribbon)
  }

  # prior SD on working scale
  df <- dplyr::mutate(df, sigma = sqrt(1 / .data$Priors))

  # transform 95% band to raw scale
  trans_q <- function(mu, s, scale) {
    if (scale == "log")        c(lo = exp(mu - 1.96*s), hi = exp(mu + 1.96*s))
    else if (scale == "log10") c(lo = 10^(mu - 1.96*s), hi = 10^(mu + 1.96*s))
    else                       c(lo = mu - 1.96*s,      hi = mu + 1.96*s)
  }

  purrr::pmap_dfr(df, function(Parameter, Priors, sigma, ...) {
    mu <- unname(mu_hyp[Parameter])
    q  <- trans_q(mu, sigma, mu_scale)
    tibble::tibble(Parameter = Parameter, Priors = Priors,
                   prior_lo = q["lo"], prior_hi = q["hi"])
  })
}
```



```{r echo=F, warning=F, notes=F, include = F}
# ---------- panel A: posterior only (blue ribbon + dashed prior mean) ----------
plot_dx_posterior <- function(mod_out,
                              param    = c("y0","y1","t1","shape","alpha"),
                              prior    = "Precision",
                              mu_hyp   = NULL,
                              mu_scale = c("log","log10","identity")[1]) {

  mu_scale <- match.arg(mu_scale)
  mod_out  <- dplyr::filter(mod_out, .data$Parameter %in% param)

  post_sum <- post_summ_by_prior(mod_out, param = param) %>%
    dplyr::transmute(
      Parameter, Priors,
      median  = .data$Median,
      two_5   = .data$`2.5%`,
      nine_75 = .data$`97.5%`
    )

  # dashed prior means
  prior_means_df <- NULL
  if (!is.null(mu_hyp)) {
    ok <- intersect(names(mu_hyp), unique(post_sum$Parameter))
    if (length(ok) > 0) {
      to_raw <- function(x) switch(mu_scale, log=exp(x), log10=10^x, identity=x)
      prior_means_df <- tibble::tibble(
        Parameter  = ok,
        prior_mean = to_raw(unname(mu_hyp[ok]))
      )
    }
  }

  g <- ggplot2::ggplot(post_sum, ggplot2::aes(x = Priors, y = median)) +
    ggplot2::geom_ribbon(ggplot2::aes(ymin = two_5, ymax = nine_75),
                         fill = "lightblue", alpha = 0.4) +
    ggplot2::geom_line() +
    ggplot2::theme_bw() +
    ggplot2::labs(
      y = "Posterior median (with 95% CI)",
      x = paste0(prior, " values"),
      caption = "Blue ribbon = 95% posterior; dashed = prior mean"
    ) +
    ggplot2::scale_x_log10() +
    ggplot2::scale_y_log10() +
    ggplot2::facet_wrap(~Parameter, scales = "free_y")

  if (!is.null(prior_means_df)) {
    g <- g + ggplot2::geom_hline(data = prior_means_df,
                                 ggplot2::aes(yintercept = prior_mean),
                                 linetype = "dashed", color = "black")
  }
  g
}
```



```{r echo=F, warning=F, notes=F, include = F}
# ---------- panel B: prior only (grey ribbon + dashed prior mean) ----------
plot_dx_prior <- function(mod_out,
                          param    = c("y0","y1","t1","shape","alpha"),
                          prior    = "Precision",
                          mu_hyp   = NULL,
                          mu_scale = c("log","log10","identity")[1],
                          p_min_for_ribbon = NULL) {

  mu_scale <- match.arg(mu_scale)
  mod_out  <- dplyr::filter(mod_out, .data$Parameter %in% param)

  # we just need the {Parameter, Priors} grid (use posterior summary to get it cleanly)
  post_sum <- post_summ_by_prior(mod_out, param = param) %>%
    dplyr::distinct(Parameter, Priors)

  prior_ribbon_df <- .make_prior_ribbon(post_sum, mu_hyp, mu_scale, p_min_for_ribbon)

  # dashed prior means
  prior_means_df <- NULL
  if (!is.null(mu_hyp)) {
    ok <- intersect(names(mu_hyp), unique(post_sum$Parameter))
    if (length(ok) > 0) {
      to_raw <- function(x) switch(mu_scale, log=exp(x), log10=10^x, identity=x)
      prior_means_df <- tibble::tibble(
        Parameter  = ok,
        prior_mean = to_raw(unname(mu_hyp[ok]))
      )
    }
  }

  ggplot2::ggplot(prior_ribbon_df, ggplot2::aes(x = Priors)) +
    ggplot2::geom_ribbon(ggplot2::aes(ymin = prior_lo, ymax = prior_hi),
                         fill = "grey70", alpha = 0.35) +
    ggplot2::theme_bw() +
    ggplot2::labs(
      y = "95% prior range",
      x = paste0(prior, " values"),
      caption = paste0(
        "Grey ribbon = 95% prior; dashed = prior mean",
        if (!is.null(p_min_for_ribbon)) paste0(" (ribbon drawn for Priors ≥ ", p_min_for_ribbon, ")")
      )
    ) +
    ggplot2::scale_x_log10() +
    ggplot2::scale_y_log10() +
    ggplot2::facet_wrap(~Parameter, scales = "free_y") +
    { if (!is.null(prior_means_df))
        ggplot2::geom_hline(data = prior_means_df,
                            ggplot2::aes(yintercept = prior_mean),
                            linetype = "dashed", color = "black") }
}
```


```{r echo=F, warning=F, notes=F, include = F}
# assumes you already have post_summ_by_prior() defined
plot_dx_two_row <- function(mod_out,
                            param    = c("y0","y1","t1","shape","alpha"),
                            prior    = "Precision",
                            mu_hyp   = NULL,                    # named vector of prior means
                            mu_scale = c("log","log10","identity")[1],
                            p_min_for_ribbon = NULL) {          # e.g., 0.01 to trim huge prior bands
  mu_scale <- match.arg(mu_scale)

  mod_out  <- dplyr::filter(mod_out, .data$Parameter %in% param)

  # --- posterior summary on raw plotting scale
  post_sum <- post_summ_by_prior(mod_out, param = param) %>%
    dplyr::transmute(
      Parameter, Priors,
      median  = .data$Median,
      two_5   = .data$`2.5%`,
      nine_75 = .data$`97.5%`
    )

  # --- dashed prior means (on raw scale)
  to_raw <- function(x) switch(mu_scale, log=exp(x), log10=10^x, identity=x)

  prior_means_df <- NULL
  if (!is.null(mu_hyp)) {
    ok <- intersect(names(mu_hyp), unique(post_sum$Parameter))
    if (length(ok) > 0) {
      prior_means_df <- tibble::tibble(
        Parameter  = ok,
        prior_mean = to_raw(unname(mu_hyp[ok]))
      ) |>
        tidyr::expand_grid(panel = c("Posterior","Prior"))   # draw line in both rows
    }
  }

  # --- build 95% PRIOR ribbon on raw scale; optional clipping of tiny precisions
  make_prior_ribbon <- function(df_params_priors, mu_hyp, mu_scale, pmin) {
    if (is.null(mu_hyp)) return(NULL)
    keep <- intersect(df_params_priors$Parameter, names(mu_hyp))
    if (length(keep) == 0) return(NULL)

    df <- dplyr::filter(df_params_priors, .data$Parameter %in% keep)
    if (!is.null(pmin)) df <- dplyr::filter(df, .data$Priors >= pmin)

    df <- dplyr::mutate(df, sigma = sqrt(1 / .data$Priors))
    trans_q <- function(mu, s, scale) {
      if (scale == "log")        c(lo = exp(mu - 1.96*s), hi = exp(mu + 1.96*s))
      else if (scale == "log10") c(lo = 10^(mu - 1.96*s), hi = 10^(mu + 1.96*s))
      else                       c(lo = mu - 1.96*s,      hi = mu + 1.96*s)
    }
    purrr::pmap_dfr(df, function(Parameter, Priors, sigma, ...) {
      mu <- unname(mu_hyp[Parameter]); q <- trans_q(mu, sigma, mu_scale)
      tibble::tibble(Parameter = Parameter, Priors = Priors,
                     prior_lo = q["lo"], prior_hi = q["hi"])
    })
  }

  prior_grid        <- dplyr::distinct(post_sum, Parameter, Priors)
  prior_ribbon_df   <- make_prior_ribbon(prior_grid, mu_hyp, mu_scale, p_min_for_ribbon)

  # tag data with "panel" for facet_grid(panel ~ Parameter)
  post_sum$panel      <- "Posterior"
  if (!is.null(prior_ribbon_df)) prior_ribbon_df$panel <- "Prior"

  g <- ggplot2::ggplot(mapping = ggplot2::aes(x = Priors)) +
    # PRIOR row (grey ribbon)
    { if (!is.null(prior_ribbon_df))
        ggplot2::geom_ribbon(data = prior_ribbon_df,
                             ggplot2::aes(ymin = prior_lo, ymax = prior_hi),
                             fill = "grey70", alpha = 0.35) } +
    # POSTERIOR row (blue ribbon + median line)
    ggplot2::geom_ribbon(data = post_sum,
                         ggplot2::aes(ymin = two_5, ymax = nine_75),
                         fill = "lightblue", alpha = 0.4) +
    ggplot2::geom_line(data = post_sum, ggplot2::aes(y = median)) +
    # dashed prior mean (both rows)
    { if (!is.null(prior_means_df))
        ggplot2::geom_hline(data = prior_means_df,
                            ggplot2::aes(yintercept = prior_mean),
                            linetype = "dashed", color = "black") } +
    ggplot2::theme_bw() +
    ggplot2::labs(
  x = "Precision values",
  caption = "Top row: Posterior (blue 95% CI + median).\nBottom row: Prior (grey 95% range).\nDashed = prior mean."
) +
    ggplot2::scale_x_log10() +
    ggplot2::scale_y_log10() +
    ggplot2::facet_grid(rows = ggplot2::vars(panel),
                        cols  = ggplot2::vars(Parameter),
                        scales = "free_y") +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "grey95"),
      panel.spacing.y  = grid::unit(0.6, "lines"),
      panel.spacing.x  = grid::unit(0.8, "lines")
    )

  g
}

```


```{r echo=F, warning=F, notes=F, message = F}
# Iterating over different prior values for shigella

load(here::here() %>% fs::path("/data/shigella_y0.rda"))
load(here::here() %>% fs::path("/data/shigella_y1.rda"))
load(here::here() %>% fs::path("/data/shigella_t.rda"))
load(here::here() %>% fs::path("/data/shigella_r.rda"))
load(here::here() %>% fs::path("/data/shigella_alpha.rda"))

```


------------------------------------------------------------------------

## Shigella Precision Modeling

### Initial antibody level (y0)

Shigella IpaB-IgG Rhat values by varying `prec_hyp_param` prior values for y0

```{r echo=F, warning=F, notes=F, include = F}
plot_dx_posterior(shigella_y0, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```


```{r echo=F, warning=F, notes=F, include = F}
plot_dx_prior(shigella_y0, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```


```{r echo=F, warning=F, notes=F, include = F}
plot_dx_two_row(
  shigella_y0,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```

Across the precision sweep, $t_{1}$ (and to a lesser extent $y_{0}$) remain near their prior means, suggesting prior–data agreement or some prior influence, 
whereas $\alpha$, $r$ (shape), and $y_{1}$ are pulled away from their prior means, indicating the data are informative and those prior means may be miscentered; prior ribbons and modern diagnostics confirm these impressions.


------------------------------------------------------------------------

### Peak antibody concentration (y1)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for y1

```{r echo=F, warning=F, notes=F, include = F}
plot_dx_posterior(shigella_y1, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```


```{r echo=F, warning=F, notes=F, include = F}
plot_dx_prior(shigella_y1, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```

```{r echo=F, warning=F, notes=F, include = F}
plot_dx_two_row(
  shigella_y1,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```


------------------------------------------------------------------------

### Time to peak (t1)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for t1

```{r echo=F, warning=F, notes=F, include = F}
plot_dx_posterior(shigella_t, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```

```{r echo=F, warning=F, notes=F, include = F}
plot_dx_prior(shigella_t, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```


```{r echo=F, warning=F, notes=F, include = F}
plot_dx_two_row(
  shigella_t,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```


------------------------------------------------------------------------

### Shape of decay (r)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for r

```{r echo=F, warning=F, notes=F, include = F}
plot_dx_posterior(shigella_r, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```


```{r echo=F, warning=F, notes=F, include = F}
plot_dx_prior(shigella_r, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```


```{r echo=F, warning=F, notes=F, include = F}
plot_dx_two_row(
  shigella_r,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```


------------------------------------------------------------------------

### Decay rate (alpha)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for alpha

```{r echo=F, warning=F}
plot_dx_posterior(shigella_alpha, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```

```{r echo=F, warning=F}
plot_dx_prior(shigella_alpha, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```


```{r echo=F, warning=F, fig.width=10,fig.height=5}
plot_dx_two_row(
  shigella_alpha,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```



------------------------------------------------------------------------

# Conclusion

