---
title: "Priors Analysis of Shigella"
subtitle: "Test 1: Sensitivity Analysis of Prior Precision for Shigella (prec_hyp_param)"
author: "Kwan Ho Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true             # enable table of contents
    toc_depth: 3          # up to three heading levels (#, ##, ###)
    toc_float: true       # floating TOC on the side
    highlight: tango      # syntax highlighting style
  pdf_document:
    toc: true             # also enable table of contents for PDF
    toc_depth: 3          # same heading depth for PDF
    highlight: tango      # consistent highlighting
---

```{r echo=F, include=F, warning=F, notes = F}

library(dplyr)
library(runjags)
library(readr)
library(serodynamics)
library(serocalculator)
library(tidyverse)
library(runjags)
library(coda)
library(ggmcmc)
library(ggbeeswarm)
library(ggpubr)
library(knitr)
library(table1)
library(readxl)
devtools::load_all()

options(scipen = 99, digits = 4)

#model file
file.mod <- fs::path_package("serodynamics", "extdata/model.jags")
```

```{r}
#| label: "load shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'
df <- 
  fs::path_package(
    package = "shigella",
    "extdata/3.8.2024 Compiled Shigella datav2.xlsx"
  ) |> 
  read_excel(sheet = "Compiled")
```

```{r}
#| label: "manipulate Shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

# filter ipab
dL_clean_Ipab <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_ipab_MFI
)

dL_clean_Ipab_IgG<-dL_clean_Ipab %>%
  filter(antigen_iso == "IgG")

```

# Analaysis overview:

In this analysis, we are examining how **changing hyperpriors** for

-   `prec_hyp_param` (spread/uncertainty) and

-   `mu_hyp_param` (center/average guess)

affects **R̂ values (convergence)** and **parameter estimates** in our Bayesian hierarchical model.

The Shigella dataset is filtered to include only the **IpaB_IgG biomarker**. All JAGS models were run with the following constants:

-   `nchain = 4`, `nadapt = 1000`, `nburn = 500`, `nmc = 2500`, `niter = 5000`.

# Test 1: Sensitivity analysis of prior spread (`prec_hyp_param`)

**Goal**: Check how the model behaves when we change the **prior precision** (spread) for each parameter, one at a time. This tests whether the model is sensitive to overly tight or overly vague priors, and helps identify a safe level of spread that supports good convergence.

**Setup**:

-   Default priors for means:

    `mu.hyp = c(1.0, 7.0, 1.0, -2.0, -3.0)`

-   Precision sweep values:

    `prec.hyp = {10000, 1000, 100, 10, 1, 0.1, 0.01, 0.001, 0.00001}`

-   Other hyperpriors:

    `omega = diag(c(10,10,10,10,10,10)), wishdf = 20, prec.logy.hyp = c(4.0,1)`

**Method**:

-   Vary the precision for one parameter at a time (e.g., $y_0$), while holding the other four precisions fixed.

-   For each run, record how $\hat{R}$ and parameter estimates for all parameters respond.

**Interpretation**:

-   High precision (narrow prior) forces estimates close to the mean guess.

-   Low precision (diffuse prior) gives more freedom but may cause poor convergence.

-   The goal is to identify a stable range of precision values that allow good convergence ($\hat{R} \approx 1$, stable ESS) without over-constraining the model.

------------------------------------------------------------------------

```{r echo=F, warning=F, notes=F, include = F}
# --- summarize with serodynamics::post_summ() by prior value -----------------
post_summ_by_prior <- function(data,
                               iso   = unique(data$Iso_type),
                               param = unique(data$Parameter),
                               strat = unique(data$Stratification)) {
  # runs post_summ() separately for each Priors value and binds results
  data %>%
    dplyr::group_by(Priors) %>%
    dplyr::group_modify(~ serodynamics::post_summ(.x, iso = iso,
                                                  param = param, strat = strat)) %>%
    dplyr::ungroup()
}

mu_vec <- c(y0 = 1.0, y1 = 7.0, t1 = 1.0, alpha = -2.0, shape = -3.0)  # from setup
```

```{r echo=F, warning=F, notes=F, include = F}
# ---- main plotting function: type = "estimate" or "rhat" -------------------
plot_dx <- function(mod_out,
                    type  = c("estimate","rhat")[1],
                    param = c("y0","y1","t1","shape","alpha"),
                    prior = "Precision",
                    mu_hyp  = NULL,
                    mu_scale = c("log","log10","identity")[1]) {
  
  type <- tolower(type)
  if (!type %in% c("estimate","rhat")) stop("`type` must be 'estimate' or 'rhat'.")
  
  mod_out <- mod_out %>% dplyr::filter(.data$Parameter %in% param)
  
  to_raw <- function(x, scale) {
    switch(scale,
           "log"      = exp(x),
           "log10"    = 10^x,
           "identity" = x)
  }
  
  if (type == "estimate") {
    post_sum <- post_summ_by_prior(mod_out, param = param) %>%
      dplyr::transmute(Parameter, Priors,
                       median  = .data$Median,
                       two_5   = .data$`2.5%`,
                       nine_75 = .data$`97.5%`)
    
    prior_means_df <- NULL
    if (!is.null(mu_hyp)) {
      ok <- intersect(names(mu_hyp), unique(post_sum$Parameter))
      if (length(ok) > 0) {
        prior_means_df <- tibble::tibble(
          Parameter  = ok,
          prior_mean = to_raw(unname(mu_hyp[ok]), mu_scale)
        )
      }
    }
    
    g <- ggplot2::ggplot(post_sum, ggplot2::aes(x = Priors, y = median)) +
      ggplot2::geom_line() +
      ggplot2::geom_ribbon(ggplot2::aes(ymin = two_5, ymax = nine_75),
                           fill = "lightblue", alpha = 0.4) +
      ggplot2::theme_bw() +
      ggplot2::labs(y = "Posterior median (with 95% CI)",
                    x = paste0(prior, " values"),
                    caption = if (!is.null(prior_means_df))
                      paste0("Dashed lines = prior means (mu_scale = '", mu_scale, "').")) +
      ggplot2::scale_x_log10() +
      ggplot2::scale_y_log10() +
      ggplot2::facet_wrap(~Parameter, scales = "free_y")
    
    if (!is.null(prior_means_df)) {
      g <- g + ggplot2::geom_hline(data = prior_means_df,
                                   ggplot2::aes(yintercept = prior_mean),
                                   linetype = "dashed", color = "black")
    }
    return(g)
  }
  
  # -------- rhat branch --------
  Rhats <- data.frame(Parameter = character(), Rhat = numeric(), Priors = numeric())
  
  for (i in sort(unique(mod_out$Priors))) {
    df_i <- mod_out %>% dplyr::filter(.data$Priors == i)
    
    # sanity check: required columns
    need <- c("Iteration","Chain","Parameter","value")
    if (!all(need %in% names(df_i))) {
      stop("mod_out must contain columns: ", paste(need, collapse = ", "))
    }
    
    rhat_i <- ggmcmc::ggs_Rhat(df_i, plot = FALSE) %>%
      dplyr::mutate(Priors = i)
    
    Rhats <- dplyr::bind_rows(Rhats, rhat_i)
  }
  
  ggplot2::ggplot(Rhats,
                  ggplot2::aes(x = Priors, y = Rhat,
                               color = Parameter, group = Parameter)) +
    ggplot2::geom_line() +
    ggplot2::geom_point(size = 1.2) +
    ggplot2::geom_hline(yintercept = 1.10, color = "red",       linetype = "dashed") +
    ggplot2::geom_hline(yintercept = 1.05, color = "darkorange", linetype = "dotted") +
    ggplot2::theme_bw() +
    ggplot2::labs(y = expression(hat(R)),
                  x = paste0(prior, " values"),
                  caption = "Dashed red = 1.10; dotted orange = 1.05") +
    ggplot2::scale_x_log10(limits = c(min(Rhats$Priors), max(Rhats$Priors))) +
    ggplot2::scale_color_brewer(palette = "Paired")
}

```

```{r echo=F, warning=F, notes=F, include = F}
# ---------- helper: 95% prior ribbon on the raw plotting scale ----------
.make_prior_ribbon <- function(df_params_priors, mu_hyp, mu_scale = c("log","log10","identity")[1],
                               p_min_for_ribbon = NULL) {
  if (is.null(mu_hyp)) return(NULL)
  mu_scale <- match.arg(mu_scale)

  # df_params_priors must have: Parameter, Priors (precision grid)
  keep <- intersect(df_params_priors$Parameter, names(mu_hyp))
  if (length(keep) == 0) return(NULL)

  df <- dplyr::filter(df_params_priors, .data$Parameter %in% keep)

  # optional clipping of tiny precisions (helps avoid absurd ranges)
  if (!is.null(p_min_for_ribbon)) {
    df <- dplyr::filter(df, .data$Priors >= p_min_for_ribbon)
  }

  # prior SD on working scale
  df <- dplyr::mutate(df, sigma = sqrt(1 / .data$Priors))

  # transform 95% band to raw scale
  trans_q <- function(mu, s, scale) {
    if (scale == "log")        c(lo = exp(mu - 1.96*s), hi = exp(mu + 1.96*s))
    else if (scale == "log10") c(lo = 10^(mu - 1.96*s), hi = 10^(mu + 1.96*s))
    else                       c(lo = mu - 1.96*s,      hi = mu + 1.96*s)
  }

  purrr::pmap_dfr(df, function(Parameter, Priors, sigma, ...) {
    mu <- unname(mu_hyp[Parameter])
    q  <- trans_q(mu, sigma, mu_scale)
    tibble::tibble(Parameter = Parameter, Priors = Priors,
                   prior_lo = q["lo"], prior_hi = q["hi"])
  })
}
```

```{r echo=F, warning=F, notes=F, include = F}
# ---------- panel A: posterior only (blue ribbon + dashed prior mean) ----------
plot_dx_posterior <- function(mod_out,
                              param    = c("y0","y1","t1","shape","alpha"),
                              prior    = "Precision",
                              mu_hyp   = NULL,
                              mu_scale = c("log","log10","identity")[1]) {

  mu_scale <- match.arg(mu_scale)
  mod_out  <- dplyr::filter(mod_out, .data$Parameter %in% param)

  post_sum <- post_summ_by_prior(mod_out, param = param) %>%
    dplyr::transmute(
      Parameter, Priors,
      median  = .data$Median,
      two_5   = .data$`2.5%`,
      nine_75 = .data$`97.5%`
    )

  # dashed prior means
  prior_means_df <- NULL
  if (!is.null(mu_hyp)) {
    ok <- intersect(names(mu_hyp), unique(post_sum$Parameter))
    if (length(ok) > 0) {
      to_raw <- function(x) switch(mu_scale, log=exp(x), log10=10^x, identity=x)
      prior_means_df <- tibble::tibble(
        Parameter  = ok,
        prior_mean = to_raw(unname(mu_hyp[ok]))
      )
    }
  }

  g <- ggplot2::ggplot(post_sum, ggplot2::aes(x = Priors, y = median)) +
    ggplot2::geom_ribbon(ggplot2::aes(ymin = two_5, ymax = nine_75),
                         fill = "lightblue", alpha = 0.4) +
    ggplot2::geom_line() +
    ggplot2::theme_bw() +
    ggplot2::labs(
      y = "Posterior median (with 95% CI)",
      x = paste0(prior, " values"),
      caption = "Blue ribbon = 95% posterior; dashed = prior mean"
    ) +
    ggplot2::scale_x_log10() +
    ggplot2::scale_y_log10() +
    ggplot2::facet_wrap(~Parameter, scales = "free_y")

  if (!is.null(prior_means_df)) {
    g <- g + ggplot2::geom_hline(data = prior_means_df,
                                 ggplot2::aes(yintercept = prior_mean),
                                 linetype = "dashed", color = "black")
  }
  g
}
```

```{r echo=F, warning=F, notes=F, include = F}
# ---------- panel B: prior only (grey ribbon + dashed prior mean) ----------
plot_dx_prior <- function(mod_out,
                          param    = c("y0","y1","t1","shape","alpha"),
                          prior    = "Precision",
                          mu_hyp   = NULL,
                          mu_scale = c("log","log10","identity")[1],
                          p_min_for_ribbon = NULL) {

  mu_scale <- match.arg(mu_scale)
  mod_out  <- dplyr::filter(mod_out, .data$Parameter %in% param)

  # we just need the {Parameter, Priors} grid (use posterior summary to get it cleanly)
  post_sum <- post_summ_by_prior(mod_out, param = param) %>%
    dplyr::distinct(Parameter, Priors)

  prior_ribbon_df <- .make_prior_ribbon(post_sum, mu_hyp, mu_scale, p_min_for_ribbon)

  # dashed prior means
  prior_means_df <- NULL
  if (!is.null(mu_hyp)) {
    ok <- intersect(names(mu_hyp), unique(post_sum$Parameter))
    if (length(ok) > 0) {
      to_raw <- function(x) switch(mu_scale, log=exp(x), log10=10^x, identity=x)
      prior_means_df <- tibble::tibble(
        Parameter  = ok,
        prior_mean = to_raw(unname(mu_hyp[ok]))
      )
    }
  }

  ggplot2::ggplot(prior_ribbon_df, ggplot2::aes(x = Priors)) +
    ggplot2::geom_ribbon(ggplot2::aes(ymin = prior_lo, ymax = prior_hi),
                         fill = "grey70", alpha = 0.35) +
    ggplot2::theme_bw() +
    ggplot2::labs(
      y = "95% prior range",
      x = paste0(prior, " values"),
      caption = paste0(
        "Grey ribbon = 95% prior; dashed = prior mean",
        if (!is.null(p_min_for_ribbon)) paste0(" (ribbon drawn for Priors ≥ ", p_min_for_ribbon, ")")
      )
    ) +
    ggplot2::scale_x_log10() +
    ggplot2::scale_y_log10() +
    ggplot2::facet_wrap(~Parameter, scales = "free_y") +
    { if (!is.null(prior_means_df))
        ggplot2::geom_hline(data = prior_means_df,
                            ggplot2::aes(yintercept = prior_mean),
                            linetype = "dashed", color = "black") }
}
```

```{r echo=F, warning=F, notes=F, include = F}
# assumes we already have post_summ_by_prior() defined
plot_dx_two_row <- function(mod_out,
                            param    = c("y0","y1","t1","shape","alpha"),
                            prior    = "Precision",
                            mu_hyp   = NULL,                    # named vector of prior means
                            mu_scale = c("log","log10","identity")[1],
                            p_min_for_ribbon = NULL) {          # e.g., 0.01 to trim huge prior bands
  mu_scale <- match.arg(mu_scale)

  mod_out  <- dplyr::filter(mod_out, .data$Parameter %in% param)

  # --- posterior summary on raw plotting scale
  post_sum <- post_summ_by_prior(mod_out, param = param) %>%
    dplyr::transmute(
      Parameter, Priors,
      median  = .data$Median,
      two_5   = .data$`2.5%`,
      nine_75 = .data$`97.5%`
    )

  # --- dashed prior means (on raw scale)
  to_raw <- function(x) switch(mu_scale, log=exp(x), log10=10^x, identity=x)

  prior_means_df <- NULL
  if (!is.null(mu_hyp)) {
    ok <- intersect(names(mu_hyp), unique(post_sum$Parameter))
    if (length(ok) > 0) {
      prior_means_df <- tibble::tibble(
        Parameter  = ok,
        prior_mean = to_raw(unname(mu_hyp[ok]))
      ) |>
        tidyr::expand_grid(panel = c("Posterior","Prior"))   # draw line in both rows
    }
  }

  # --- build 95% PRIOR ribbon on raw scale; optional clipping of tiny precisions
  make_prior_ribbon <- function(df_params_priors, mu_hyp, mu_scale, pmin) {
    if (is.null(mu_hyp)) return(NULL)
    keep <- intersect(df_params_priors$Parameter, names(mu_hyp))
    if (length(keep) == 0) return(NULL)

    df <- dplyr::filter(df_params_priors, .data$Parameter %in% keep)
    if (!is.null(pmin)) df <- dplyr::filter(df, .data$Priors >= pmin)

    df <- dplyr::mutate(df, sigma = sqrt(1 / .data$Priors))
    trans_q <- function(mu, s, scale) {
      if (scale == "log")        c(lo = exp(mu - 1.96*s), hi = exp(mu + 1.96*s))
      else if (scale == "log10") c(lo = 10^(mu - 1.96*s), hi = 10^(mu + 1.96*s))
      else                       c(lo = mu - 1.96*s,      hi = mu + 1.96*s)
    }
    purrr::pmap_dfr(df, function(Parameter, Priors, sigma, ...) {
      mu <- unname(mu_hyp[Parameter]); q <- trans_q(mu, sigma, mu_scale)
      tibble::tibble(Parameter = Parameter, Priors = Priors,
                     prior_lo = q["lo"], prior_hi = q["hi"])
    })
  }

  prior_grid        <- dplyr::distinct(post_sum, Parameter, Priors)
  prior_ribbon_df   <- make_prior_ribbon(prior_grid, mu_hyp, mu_scale, p_min_for_ribbon)

  # tag data with "panel" for facet_grid(panel ~ Parameter)
  post_sum$panel      <- "Posterior"
  if (!is.null(prior_ribbon_df)) prior_ribbon_df$panel <- "Prior"

  g <- ggplot2::ggplot(mapping = ggplot2::aes(x = Priors)) +
    # PRIOR row (grey ribbon)
    { if (!is.null(prior_ribbon_df))
        ggplot2::geom_ribbon(data = prior_ribbon_df,
                             ggplot2::aes(ymin = prior_lo, ymax = prior_hi),
                             fill = "grey70", alpha = 0.35) } +
    # POSTERIOR row (blue ribbon + median line)
    ggplot2::geom_ribbon(data = post_sum,
                         ggplot2::aes(ymin = two_5, ymax = nine_75),
                         fill = "lightblue", alpha = 0.4) +
    ggplot2::geom_line(data = post_sum, ggplot2::aes(y = median)) +
    # dashed prior mean (both rows)
    { if (!is.null(prior_means_df))
        ggplot2::geom_hline(data = prior_means_df,
                            ggplot2::aes(yintercept = prior_mean),
                            linetype = "dashed", color = "black") } +
    ggplot2::theme_bw() +
    ggplot2::labs(
  x = "Precision values",
  caption = "Top row: Posterior (blue 95% CI + median).\nBottom row: Prior (grey 95% range).\nDashed = prior mean."
) +
    ggplot2::scale_x_log10() +
    ggplot2::scale_y_log10() +
    ggplot2::facet_grid(rows = ggplot2::vars(panel),
                        cols  = ggplot2::vars(Parameter),
                        scales = "free_y") +
    ggplot2::theme(
      strip.background = ggplot2::element_rect(fill = "grey95"),
      panel.spacing.y  = grid::unit(0.6, "lines"),
      panel.spacing.x  = grid::unit(0.8, "lines")
    )

  g
}

```

```{r echo=F, warning=F, notes=F, message = F}
# Trace plots for a parameter across multiple prior precisions
plot_traces <- function(mod_out,
                        params,                # e.g., "y1" or c("y1","y0")
                        prior_vals,            # e.g., c(10, 1, 0.1, 0.01)
                        subject = NULL,        # e.g., "newperson"
                        ncol = NULL,           # layout columns
                        log_y = TRUE,          # log y-axis?
                        show_prior_overlay = FALSE,
                        mu_hyp = NULL,         # named vector, e.g., c(y0=1, y1=7, t1=1, shape=-3, alpha=-2)
                        mu_scale = c("log","log10","identity")[1]) {

  # --- basic checks
  stopifnot(all(c("Iteration","Chain","Parameter","Priors","value") %in% names(mod_out)))
  mu_scale <- match.arg(mu_scale)

  df <- mod_out
  if (!is.null(subject) && "Subject" %in% names(df)) {
    df <- dplyr::filter(df, .data$Subject == subject)
  }

  # allow multiple params and multiple prior values
  df <- df |>
    dplyr::filter(.data$Parameter %in% params,
                  .data$Priors %in% prior_vals) |>   # <- CHANGED: %in% to accept vectors
    dplyr::mutate(
      Chain = factor(.data$Chain),
      Priors = factor(.data$Priors, levels = sort(unique(prior_vals)))  # facet order
    )

  if (nrow(df) == 0L) {
    stop("No rows found for those params/prior_vals (check names and `Priors` present).")
  }

  # helper: transform from prior scale to raw plotting scale
  to_raw <- function(x) {
    switch(mu_scale,
           "log"      = exp(x),
           "log10"    = 10^x,
           "identity" = x)
  }

  # (optional) prior overlay (mean + 95% ribbon) per prior precision and parameter
  prior_df <- NULL
  if (isTRUE(show_prior_overlay)) {
    if (is.null(mu_hyp)) {
      stop("To use `show_prior_overlay = TRUE`, supply `mu_hyp` as a named vector (y0,y1,t1,shape,alpha).")
    }
    # build a grid of (Parameter, Priors) with mean & 95% endpoints on the raw scale
    prior_df <- tidyr::expand_grid(
      Parameter = intersect(unique(df$Parameter), names(mu_hyp)),
      Priors    = levels(df$Priors)
    ) |>
      dplyr::mutate(
        tau   = as.numeric(as.character(Priors)),
        mu    = unname(mu_hyp[Parameter]),
        sd    = 1/sqrt(tau),
        lo    = to_raw(mu - 1.96*sd),
        hi    = to_raw(mu + 1.96*sd),
        mean_ = to_raw(mu)
      )
  }

  g <- ggplot2::ggplot(df,
                       ggplot2::aes(x = .data$Iteration, y = .data$value,
                                    color = .data$Chain, group = .data$Chain)) +
    ggplot2::geom_line(alpha = 0.75, linewidth = 0.3) +
    ggplot2::facet_grid(Parameter ~ Priors, scales = if (log_y) "free_y" else "free_y") +
    ggplot2::theme_bw() +
    ggplot2::labs(x = "Iteration", y = "Sampled value", color = "Chain",
                  title = "Trace plots across prior precisions")

  if (log_y) g <- g + ggplot2::scale_y_log10()

  if (!is.null(prior_df)) {
    g <- g +
      ggplot2::geom_hline(data = prior_df,
                          ggplot2::aes(yintercept = mean_),
                          inherit.aes = FALSE, linetype = "dashed", color = "black") +
      ggplot2::geom_rect(data = prior_df,
                         ggplot2::aes(xmin = -Inf, xmax = Inf, ymin = lo, ymax = hi),
                         inherit.aes = FALSE, alpha = 0.10, fill = "grey50")
  }

  g
}



```

```{r echo=F, warning=F, notes=F, message = F}
# Iterating over different prior values for shigella

load(here::here() %>% fs::path("/data/shigella_y0_3.rda"))
load(here::here() %>% fs::path("/data/shigella_y1_3.rda"))
load(here::here() %>% fs::path("/data/shigella_t_3.rda"))
load(here::here() %>% fs::path("/data/shigella_r_3.rda"))
load(here::here() %>% fs::path("/data/shigella_alpha_3.rda"))

```

## R-hat Definition and Trace Plots

### Mathematical Definition of R-hat (Gelman–Rubin statistic)

Suppose we have $m$ Markov chains (e.g., $m=4$), each with $n$ post–burn-in samples for a parameter $\theta$.

The Gelman–Rubin statistic ($\hat{R}$) compares **within-chain** and **between-chain** variance across $m$ MCMC chains, each with $n$ samples.

1.  **Chain means**:\
    $$
    \bar{\theta}_{\cdot j} = \frac{1}{n}\sum_{i=1}^n \theta_{ij}, 
    \quad j = 1,\dots,m
    $$

2.  **Overall mean**:\
    $$
    \bar{\theta}_{\cdot \cdot} = \frac{1}{m}\sum_{j=1}^m \bar{\theta}_{\cdot j}
    $$

3.  **Between-chain variance**:\
    $$
    B = \frac{n}{m-1} \sum_{j=1}^m 
    \left( \bar{\theta}_{\cdot j} - \bar{\theta}_{\cdot \cdot} \right)^2
    $$

4.  **Within-chain variance**:\
    $$
    W = \frac{1}{m}\sum_{j=1}^m 
    \left[ \frac{1}{n-1} \sum_{i=1}^n 
    \left(\theta_{ij} - \bar{\theta}_{\cdot j}\right)^2 \right]
    $$

5.  **Marginal posterior variance estimate**:\
    $$
    \hat{\text{Var}}^+(\theta \mid y) = 
    \frac{n-1}{n}W + \frac{1}{n}B
    $$

6.  **Potential scale reduction factor (R-hat)**:\
    $$
    \hat{R} = \sqrt{\frac{\hat{\text{Var}}^+(\theta \mid y)}{W}}
    $$

**Interpretation**:\
- If $\hat{R} \approx 1$: chains mixed well → convergence likely.\
- If $\hat{R} > 1.1$: convergence questionable.\
- Larger $\hat{R}$ means the **between-chain variance (**$B$) is much larger than the **within-chain variance (**$W$).

------------------------------------------------------------------------

### Trace Plots

The trace plot is simply the sampled values $\theta_{ij}$ against iteration $i$, for each chain $j$:

$$
\text{Trace plot: } \{ (i, \theta_{ij}) : i = 1,\dots,n;\; j = 1,\dots,m \}
$$

-   If chains overlap and mix well: $B \approx W$, so $\hat{R} \to 1$.\
-   If chains are stuck in different regions or mix slowly: $B \gg W$, so $\hat{R}$ increases.

Thus, trace plots provide the raw evidence explaining why $\hat{R}$ values are high or low.

------------------------------------------------------------------------

### Application to Our Analysis

In our prior sensitivity analysis plot (R-hat values across different prior precisions):

-   Dashed red line indicates a cutoff at $\hat{R} = 1.1$.\
-   Parameters $y_0$ and *shape* spike above 1.3–1.4 at certain prior settings.\
-   This suggests possible non-convergence.

As Ezra emphasized:\
- We cannot interpret these $\hat{R}$ spikes without also looking at the **trace plots** for $y_0$ and *shape*.\
- The trace plots will reveal whether chains are mixing poorly, stuck in different modes, or simply slow to converge.

------------------------------------------------------------------------

## Shigella Precision Modeling

### Initial antibody level--Rhats (y0)

Shigella IpaB-IgG Rhat values by varying `prec_hyp_param` prior values for y0

#### What this plot is for

-   **Purpose:** To check how well MCMC chains converge when we change the prior precision for parameter $y_0$.
-   **X-axis (Priors):** The hyper-prior precision values (from wide to tight).
-   **Y-axis (**$\hat{R}$): The convergence statistic; values near 1 mean chains agree, values $> 1.1$ suggest poor convergence.
-   **Lines/points:** Each parameter ($y_0, y_1, t_1, r, \alpha$) is tracked, even though only $y_0$’s prior is varied.
-   **Dashed lines (red/orange):** Common thresholds ($\hat{R} = 1.1, 1.05$) to mark convergence problems.

#### Intention behind making it

-   **See sensitivity:** Does tweaking the prior for $y_0$ cause chain convergence to worsen or improve for $y_0$ and other parameters?
-   **Spot patterns:** If $\hat{R}$ drifts upward as precision changes, it means the sampler struggles. If $\hat{R}$ falls only because priors are too tight, that is *false convergence* (chains look good because the prior forced them).
-   **Guide next steps:** If convergence problems show up, that is a signal to tune the **sampler settings** (step size, acceptance rate, etc.), not just tighten priors.

```{r echo=F, warning=F, notes=F, message = F}
#| label: fig-rhat-y0
plot_dx(shigella_y0_3, type = "Rhat")
```

### Traceplots (y0)

#### What these plots are for

-   **Purpose:** Visualize how MCMC chains behave across different **prior precisions** for each parameter.
-   **X-axis:** Iterations of the sampler.
-   **Y-axis:** Sampled parameter values (log scale).
-   **Facets:** Each column = a different prior precision (0.01, 0.1, 1, 100).
-   **Colored lines:** Different chains (1–4).
-   **Dashed line + gray band:** Prior mean and 95% prior interval (if overlay shown).

#### Intention behind making them

-   **Check mixing:** Do chains overlap and “wander” freely, or do they stick in different regions?
-   **Spot prior influence:** If chains collapse tightly around the prior mean (flat lines inside the band), the prior is dominating (too informative).
-   **Compare sensitivity:** Side-by-side precisions reveal how behavior changes as priors go from loose (wide) to tight (narrow).
-   **Complement** $\hat{R}$: These traces are the raw evidence behind $\hat{R}$—use them to see if a specific chain drifts or if apparent convergence is prior-induced.

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6,fig.cap="Trace plots for y0 across different prior precisions (y0 prior varied)"}
#| label: fig-traceplots-y0-y0
plot_traces(
  mod_out   = shigella_y0_3,
  params    = "y0",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6, fig.cap="Trace plots for y1 across different prior precisions while varying the y0 prior"}
#| label: fig-traceplots-y0-y1
plot_traces(
  mod_out   = shigella_y0_3,
  params    = "y1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6, fig.cap="Trace plots for t1 across different prior precisions while varying the y0 prior"}
#| label: fig-traceplots-y0-t1
plot_traces(
  mod_out   = shigella_y0_3,
  params    = "t1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6, fig.cap="Trace plots for shape parameter across different prior precisions while varying the y0 prior"}
#| label: fig-traceplots-y0-shape
plot_traces(
  mod_out   = shigella_y0_3,
  params    = "shape",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6, fig.cap="Trace plots for alpha parameter across different prior precisions while varying the y0 prior"}
#| label: fig-traceplots-y0-alpha
plot_traces(
  mod_out   = shigella_y0_3,
  params    = "alpha",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

**Convergence Diagnostics: R-hat vs Trace Plots**

The R-hat values across different parameters generally stayed close to 1, suggesting that between-chain and within-chain variances were similar and therefore the MCMC chains were not grossly divergent:

$$ \hat{R} = \sqrt{\frac{\tfrac{n-1}{n}W + \tfrac{1}{n}B}{W}} \approx 1 $$

However, the trace plots provide a more nuanced picture. While some parameters (e.g., $\alpha$) show stable mixing across chains, others (especially $y_0$ and $y_1$) display poor mixing and large excursions, indicating non-convergence despite $\hat{R}$ values being near 1.

Parameters such as $t_1$ highlight the limitation of relying solely on $\hat{R}$: even when $\hat{R} \approx 1$, the trace plots reveal high autocorrelation and noisy trajectories, showing that the chains have not fully explored the posterior efficiently.

**Interpretation:**\
- Both $\hat{R}$ and trace plots are essential: $\hat{R}$ alone can mask poor mixing if all chains share similar variance but are highly autocorrelated.

### Posterior estimates across prior precision values

#### What this plot is for

-   **Purpose:** Examine how the **posterior estimates** (median + 95% CI) for each parameter respond as we vary the **hyper-prior precision for** $y_0$.
-   **X-axis (Priors):** Precision values applied to $y_0$’s hyper-prior (very wide → very tight).
-   **Y-axis:** Posterior median with 95% credible interval (log scale).
-   **Facets:** One panel per parameter ($y_0$, $y_1$, $t_1$, shape, $\alpha$).
-   **Shading/lines:** Blue ribbon = 95% posterior CI; black dashed line = prior mean for that parameter.

#### Intention behind making it

-   **Check robustness:** Do posterior medians stay stable as prior precision changes, or do they shift a lot?
-   **Detect prior dominance:** If posteriors collapse onto the dashed prior mean, the prior is too informative.
-   **Compare parameters:** Identify which parameters are sensitive to $y_0$’s prior precision (e.g., $y_0$) and which are relatively stable (e.g., $y_1$, $t_1$).
-   **Guide prior choice:** Find a **safe precision range** where posteriors are data-driven (not prior-driven) yet estimates remain well-behaved.

```{r echo=F, warning=F, notes=F}
#| label: fig-posterior-y0
plot_dx_posterior(shigella_y0_3, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```

### Prior distributions across precision values

#### What this plot is for

-   **Purpose:** Display the **prior distributions** alone (no data), across different **hyper-prior precision** values.
-   **X-axis (Priors):** Precision values (wide → tight).
-   **Y-axis:** Parameter scale implied by the prior (log scale).
-   **Facets:** One panel per parameter ($y_0$, $y_1$, $t_1$, shape, $\alpha$).
-   **Shading/lines:** Gray ribbon = 95% prior interval; black dashed line = prior mean.

#### Intention behind making it

-   **Visualize assumptions:** Make explicit how wide/narrow each prior is for each precision setting.
-   **Check scaling:** Confirm prior ranges are sensible for the assay scale (e.g., ELISA vs. multiplex units).
-   **Diagnose informativeness:** Very narrow ribbons (high precision) imply a **strong/controlling** prior; very wide ribbons imply a **weak** prior that lets data dominate.
-   **Link to posterior:** When compared with posterior plots, reveals whether results are **prior-driven** or **data-driven**.

```{r echo=F, warning=F, notes=F}
#| label: fig-prior-y0
plot_dx_prior(shigella_y0_3, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```

### Posterior versus prior comparison

```{r echo=F, warning=F, notes=F, fig.width=12,fig.height=6}
#| label: fig-comparison-y0
plot_dx_two_row(
  shigella_y0_3,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```

------------------------------------------------------------------------

### Peak antibody concentration (y1)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for y1

```{r echo=F, warning=F, notes=F, message = F}
#| label: fig-rhat-y1
plot_dx(shigella_y1_3, type = "Rhat")
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-y1-y0
plot_traces(
  mod_out   = shigella_y1_3,
  params    = "y0",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-y1-y1
plot_traces(
  mod_out   = shigella_y1_3,
  params    = "y1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-y1-t1
plot_traces(
  mod_out   = shigella_y1_3,
  params    = "t1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-y1-shape
plot_traces(
  mod_out   = shigella_y1_3,
  params    = "shape",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-y1-alpha
plot_traces(
  mod_out   = shigella_y1_3,
  params    = "alpha",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

#### Posterior estimates across prior precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-posterior-y1
plot_dx_posterior(shigella_y1_3, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```

#### Prior distributions across precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-prior-y1
plot_dx_prior(shigella_y1_3, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```

#### Posterior versus prior comparison

```{r echo=F, warning=F, notes=F, fig.width=12,fig.height=6}
#| label: fig-comparison-y1
plot_dx_two_row(
  shigella_y1_3,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```

------------------------------------------------------------------------

### Time to peak (t1)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for t1

```{r echo=F, warning=F, notes=F, message = F}
#| label: fig-rhat-t1
plot_dx(shigella_t_3, type = "Rhat")
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-t1-y0
plot_traces(
  mod_out   = shigella_t_3,
  params    = "y0",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-t1-y1
plot_traces(
  mod_out   = shigella_t_3,
  params    = "y1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-t1-t1
plot_traces(
  mod_out   = shigella_t_3,
  params    = "t1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-t1-shape
plot_traces(
  mod_out   = shigella_t_3,
  params    = "shape",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-t1-alpha
plot_traces(
  mod_out   = shigella_t_3,
  params    = "alpha",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

#### Posterior estimates across prior precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-posterior-t1
plot_dx_posterior(shigella_t_3, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```

#### Prior distributions across precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-prior-t1
plot_dx_prior(shigella_t_3, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```

#### Posterior versus prior comparison

```{r echo=F, warning=F, notes=F, fig.width=12,fig.height=6}
#| label: fig-comparison-t1
plot_dx_two_row(
  shigella_t_3,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```

------------------------------------------------------------------------

### Shape of decay (r)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for r

```{r echo=F, warning=F, notes=F, message = F}
#| label: fig-rhat-r
plot_dx(shigella_r_3, type = "Rhat")
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-r-y0
plot_traces(
  mod_out   = shigella_r_3,
  params    = "y0",                      # or "y1"
  prior_vals= c(1000, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-r-y1
plot_traces(
  mod_out   = shigella_r_3,
  params    = "y1",                      # or "y1"
  prior_vals= c(1000, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-r-t1
plot_traces(
  mod_out   = shigella_r_3,
  params    = "t1",                      # or "y1"
  prior_vals= c(1000, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-r-shape
plot_traces(
  mod_out   = shigella_r_3,
  params    = "shape",                      # or "y1"
  prior_vals= c(1000,10, 1,0.1),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-r-alpha
plot_traces(
  mod_out   = shigella_r_3,
  params    = "alpha",                      # or "y1"
  prior_vals= c(1000, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

#### Posterior estimates across prior precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-posterior-r
plot_dx_posterior(shigella_r_3, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```

#### Prior distributions across precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-prior-r
plot_dx_prior(shigella_r_3, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```

#### Posterior versus prior comparison

```{r echo=F, warning=F, notes=F, fig.width=12,fig.height=6}
#| label: fig-comparison-r
plot_dx_two_row(
  shigella_r_3,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```

------------------------------------------------------------------------

### Decay rate (alpha)

Shigella Ipab-IgG Rhat values by varying `prec_hyp_param` prior values for alpha

```{r echo=F, warning=F, notes=F, message = F}
#| label: fig-rhat-alpha
plot_dx(shigella_alpha_3, type = "Rhat")
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-alpha-y0
plot_traces(
  mod_out   = shigella_alpha_3,
  params    = "y0",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-alpha-y1
plot_traces(
  mod_out   = shigella_alpha_3,
  params    = "y1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-alpha-t1
plot_traces(
  mod_out   = shigella_alpha_3,
  params    = "t1",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-alpha-shape
plot_traces(
  mod_out   = shigella_alpha_3,
  params    = "shape",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

```{r echo=F, warning=F, notes=F, message = F, fig.width=12,fig.height=6}
#| label: fig-traceplots-alpha-alpha
plot_traces(
  mod_out   = shigella_alpha_3,
  params    = "alpha",                      # or "y1"
  prior_vals= c(100, 1,0.1, 0.01),       # vector OK now
  subject   = "newperson",
  ncol      = 4,
  log_y     = TRUE,
  show_prior_overlay = TRUE,
  mu_hyp    = c(y0=1, y1=7, t1=1, shape=-3, alpha=-2),
  mu_scale  = "log"
)
```

#### Posterior estimates across prior precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-posterior-alpha
plot_dx_posterior(shigella_alpha_3, param = c("y0","y1","t1","shape","alpha"),
                  mu_hyp = mu_vec, mu_scale = "log")
```

#### Prior distributions across precision values

```{r echo=F, warning=F, notes=F}
#| label: fig-prior-alpha
plot_dx_prior(shigella_alpha_3, param = c("y0","y1","t1","shape","alpha"),
              mu_hyp = mu_vec, mu_scale = "log", p_min_for_ribbon = 0.01)
```

#### Posterior versus prior comparison

```{r echo=F, warning=F, notes=F, fig.width=12,fig.height=6}
#| label: fig-comparison-alpha
plot_dx_two_row(
  shigella_alpha_3,
  param = c("y0","y1","t1","shape","alpha"),
  mu_hyp = mu_vec,
  mu_scale = "log",
  p_min_for_ribbon = 0.01   # optional: clip ultra-vague precisions
)
```

------------------------------------------------------------------------

# Conclusion
