---
title: "Shigella-Kinetics"
author: "Kwan Ho Lee"
date: "`r Sys.Date()`"
format: pdf
jupyter: false
---


```{r}
#| label: setup
#| message: false
#| include: false
#| echo: false


library(serocalculator)
library(serodynamics)
library(readxl)
library(rlang)
library(dplyr)
library(rjags)

devtools::install()
devtools::load_all()
```

# Introduction

This vignette demonstrates modeling longitudinal antibody responses using hierarchical Bayesian models implemented in JAGS.


# Load the processed Shigella dataset

```{r,echo=FALSE, message=FALSE, warning=FALSE,results='hide'}
#| label: "load shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'
df <- 
  fs::path_package(
    package = "shigella",
    "extdata/3.8.2024 Compiled Shigella datav2.xlsx"
  ) |> 
  read_excel(sheet = "Compiled")
```

```{r,echo=FALSE, message=FALSE, warning=FALSE,results='hide'}
#| label: "manipulate Shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

# filter ipab
dL_clean_ipab <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_ipab_MFI
)
# filter sf2a
dL_clean_sf2a <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_sf2aospbsa_MFI
)
# filter sf3a
dL_clean_sf3a <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_sf3aospbsa_MFI
)
# filter sf6
dL_clean_sf6 <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_sf6ospbsa_MFI
)
# filter sonnei
dL_clean_sonnei <- process_shigella_data(
  data         = df,
  study_filter = "SOSAR",
  antigen      = n_sonneiospbsa_MFI
)
```

```{r}
#| label: "prep Shigella data"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

prepped_data_ipab <- prep_data(dL_clean_ipab)
prepped_data_sf2a <- prep_data(dL_clean_sf2a)
prepped_data_sf3a <- prep_data(dL_clean_sf3a)
prepped_data_sf6 <- prep_data(dL_clean_sf6)
prepped_data_sonnei <- prep_data(dL_clean_sonnei)

```

# Run the model using run_mod() from serodynamics package
```{r,echo=FALSE, message=FALSE, warning=FALSE,results='hide'}
#| label: "Run the model"
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

ipab_model<-run_mod(
  data=dL_clean_ipab,
  file_mod=fs::path_package("shigella", "inst/extdata/model.jags.r"),
  nchain = 4,
  nadapt = 1000,
  nburn = 100,
  nmc = 100,
  niter = 1000,
  strat = NA
)

sf2a_model<-run_mod(
  data=dL_clean_sf2a,
  file_mod=fs::path_package("shigella", "inst/extdata/model.jags.r"),
  nchain = 4,
  nadapt = 1000,
  nburn = 100,
  nmc = 100,
  niter = 1000,
  strat = NA
)
```


# Extract posterior summaries 

```{r}
#| label: "Extract posterior summaries"
#| echo: false
#| message: false
#| warning: false
curve_params_ipab <- ipab_model$jags.post$None|> 
  serodynamics:::postprocess_jags_output(
    ids = attr(prepped_data_ipab, "ids"),
    antigen_isos = attr(prepped_data_ipab, "antigens")
  )

curve_params_sf2a <- sf2a_model$jags.post$None|> 
  serodynamics:::postprocess_jags_output(
    ids = attr(prepped_data_sf2a, "ids"),
    antigen_isos = attr(prepped_data_sf2a, "antigens")
  )

```


# Use graph.curve.params()

## Plot curves

Keep using serocalculator function? or create new one with posterior predictive interval?
show all curves with median curve only + credible interval

```{r}
#| label: "Median predicted decay curves ipab"
#| echo: false
#| message: false
#| warning: false

graph.curve.params(curve_params_ipab, show_all_curves = TRUE)
graph.curve.params(curve_params_ipab)
```
```{r}
#| label: "Median predicted decay curves sf2a"
#| echo: false
#| message: false
#| warning: false

graph.curve.params(curve_params_sf2a, show_all_curves = TRUE)
graph.curve.params(curve_params_sf2a)

```

## Include 95% credible intervals for that median curve (just credible interval not PPI)


# Do this for each of the 10 antigen–isotype combinations


# Conclusion